% ============================================================
%  Math Mirror: A Self-Validating Byte-Level Transformer
%  for Agentic Mathematics
%
%  Author: Jianghan Zhang, Dartmouth College
%  Target: arXiv math.ML
% ============================================================
\documentclass[11pt,a4paper]{article}

% ── Packages (minimal set) ──────────────────────────────────
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{colortbl}
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,calc}
\usepackage{CJKutf8}
\usepackage{tcolorbox}
\tcbuselibrary{breakable,skins}

% ── Theorem environments ────────────────────────────────────
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}

% ── Spectral coloring (Newton's prism) ────────────────────
% 赤橙黄绿青蓝紫 = 7 rainbow colors on 7 environment types.
% Wavelength order (long → short = low energy → high energy):
%   Remark (赤 700nm)  → Definition (橙 620nm) → Lemma (黄 580nm)
%   → Proposition (绿 530nm) → Corollary (青 490nm)
%   → Theorem (蓝 450nm) → Proof (紫 400nm)
% The paper is a prism: white light in, 7 colors out.
% UV (Proof) suppresses IR (Remark) uncertainty.
\definecolor{specRed}{HTML}{DC143C}     % 赤 Remark
\definecolor{specOrange}{HTML}{FF8C00}  % 橙 Definition
\definecolor{specYellow}{HTML}{DAA520}  % 黄 Lemma
\definecolor{specGreen}{HTML}{228B22}   % 绿 Proposition
\definecolor{specCyan}{HTML}{00868B}    % 青 Corollary
\definecolor{specBlue}{HTML}{1E90FF}    % 蓝 Theorem
\definecolor{specViolet}{HTML}{7B2D8B}  % 紫 Proof
\tcolorboxenvironment{remark}{
  frame hidden, breakable, sharp corners,
  colback=specRed!4,
  boxsep=0pt, left=3pt, right=3pt, top=2pt, bottom=2pt,
}
\tcolorboxenvironment{definition}{
  frame hidden, breakable, sharp corners,
  colback=specOrange!5,
  boxsep=0pt, left=3pt, right=3pt, top=2pt, bottom=2pt,
}
\tcolorboxenvironment{lemma}{
  frame hidden, breakable, sharp corners,
  colback=specYellow!5,
  boxsep=0pt, left=3pt, right=3pt, top=2pt, bottom=2pt,
}
\tcolorboxenvironment{proposition}{
  frame hidden, breakable, sharp corners,
  colback=specGreen!4,
  boxsep=0pt, left=3pt, right=3pt, top=2pt, bottom=2pt,
}
\tcolorboxenvironment{corollary}{
  frame hidden, breakable, sharp corners,
  colback=specCyan!4,
  boxsep=0pt, left=3pt, right=3pt, top=2pt, bottom=2pt,
}
\tcolorboxenvironment{theorem}{
  frame hidden, breakable, sharp corners,
  colback=specBlue!5,
  boxsep=0pt, left=3pt, right=3pt, top=2pt, bottom=2pt,
}
\tcolorboxenvironment{proof}{
  frame hidden, breakable, sharp corners,
  colback=specViolet!3,
  boxsep=0pt, left=3pt, right=3pt, top=2pt, bottom=2pt,
}

% ── Notation ────────────────────────────────────────────────
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\cM}{\mathcal{M}}
\newcommand{\cV}{\mathcal{V}}
\newcommand{\cD}{\mathcal{D}}
\newcommand{\cL}{\mathcal{L}}
\newcommand{\cF}{\mathcal{F}}
\newcommand{\cG}{\mathcal{G}}
\newcommand{\cS}{\mathcal{S}}
\newcommand{\cT}{\mathcal{T}}
\newcommand{\cB}{\mathcal{B}}
\newcommand{\cA}{\mathcal{A}}
\newcommand{\cX}{\mathcal{X}}
\newcommand{\cP}{\mathcal{P}}
\newcommand{\cE}{\mathcal{E}}
\newcommand{\cR}{\mathcal{R}}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\norm}[1]{\lVert #1 \rVert}
\newcommand{\abs}[1]{\lvert #1 \rvert}
\newcommand{\DKL}{D_{\mathrm{KL}}}
\newcommand{\Breg}{D_\psi}
\newcommand{\softmax}{\mathrm{softmax}}
\newcommand{\Verify}{\mathrm{Verify}}
\newcommand{\Embed}{\mathrm{Embed}}
\newcommand{\Pullback}{\mathrm{Pullback}}
\newcommand{\Mirror}{\mathrm{Mirror}}
\newcommand{\ASCII}{\mathrm{ASCII}}

% ============================================================
\title{\textbf{Math Mirror: A Self-Validating Byte-Level Transformer\\
for Agentic Mathematics}}

\author{Jianghan Zhang\\
Department of Computer Science\\
Dartmouth College\\
\texttt{jianghan.zhang.gr@dartmouth.edu}}

\date{February 2026}

\begin{document}
\maketitle

% ============================================================
%  Abstract
% ============================================================
\begin{abstract}
We present \emph{Math Mirror}, a byte-level transformer architecture
trained exclusively on self-validated mathematical data.
The model operates over a vocabulary of 256 ASCII characters,
eliminating hidden tokenization and ensuring that the model's
internal representation is identical to the user-visible output.
Every training example is generated by symbolic computation
(SymPy) and verified before inclusion in the dataset,
guaranteeing that no mathematically incorrect statement
enters the training distribution.
At inference time, a three-level symbolic verification gate
rejects any output that cannot be confirmed correct.
We formalize the \emph{embed-prove-pullback} architecture,
in which a large language model (LLM) translates natural language
to mathematical ASCII at the boundary, the mirror computes
in pure mathematical space, and verified results return through
the LLM for human-readable presentation.
Finetuning is explicit and opt-in only: the model never learns
from conversations, preserving privacy by architecture.
We establish formal connections to the renormalization group (RG)
through mirror descent, and to the agent-nucleation framework
through the $V(T,B)$ protocol.
The \emph{Model Context Pipeline} (MCP) connects three servers---arXiv
source, multi-LLM court, and training orchestrator---into a closed
loop where masked theorem-proof pairs from arXiv drive training,
and SARSA checkpoint selection closes the feedback cycle.
An arXiv fill-in-the-blank curriculum extends the bootstrap
pipeline from synthetic identities to real mathematical proofs.
The core thesis is that a model trained only on verified mathematics
is sufficient for mathematical reasoning within its verified domain,
and that self-validation eliminates hallucination by construction.
\end{abstract}

\tableofcontents

% ── Rainbow legend ────────────────────────────────────────────
\bigskip
\begin{center}
\small
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{@{}
  >{\columncolor{specRed!4}}c
  >{\columncolor{specOrange!5}}c
  >{\columncolor{specYellow!5}}c
  >{\columncolor{specGreen!4}}c
  >{\columncolor{specCyan!4}}c
  >{\columncolor{specBlue!5}}c
  >{\columncolor{specViolet!3}}c
  @{}}
\textsc{Remark} &
\textsc{Definition} &
\textsc{Lemma} &
\textsc{Proposition} &
\textsc{Corollary} &
\textsc{Theorem} &
\textsc{Proof} \\[1pt]
{\scriptsize\begin{CJK}{UTF8}{gbsn}赤\end{CJK}\;700\,nm} &
{\scriptsize\begin{CJK}{UTF8}{gbsn}橙\end{CJK}\;620\,nm} &
{\scriptsize\begin{CJK}{UTF8}{gbsn}黄\end{CJK}\;580\,nm} &
{\scriptsize\begin{CJK}{UTF8}{gbsn}绿\end{CJK}\;530\,nm} &
{\scriptsize\begin{CJK}{UTF8}{gbsn}青\end{CJK}\;490\,nm} &
{\scriptsize\begin{CJK}{UTF8}{gbsn}蓝\end{CJK}\;450\,nm} &
{\scriptsize\begin{CJK}{UTF8}{gbsn}紫\end{CJK}\;400\,nm} \\
\end{tabular}
\renewcommand{\arraystretch}{1.0}

\smallskip
{\footnotesize
Background colours follow Newton's visible spectrum
(\begin{CJK}{UTF8}{gbsn}赤橙黄绿青蓝紫\end{CJK}).\\
Energy increases left to right: \emph{qualitative}
$\longrightarrow$ \emph{verified}.
See Remark~\ref{rem:spectral-workflow}.}
\end{center}
\bigskip

% ============================================================
%  Section 1: Introduction
% ============================================================
\section{Introduction}\label{sec:intro}

Large language models (LLMs) achieve remarkable performance across
diverse tasks, yet they suffer from a fundamental limitation in
mathematical reasoning: \emph{hallucination}~\cite{Ji2023}.
An LLM can fluently produce mathematical text that is syntactically
correct but semantically wrong---an integral with the wrong sign,
a proof with a gap, an identity that fails at a single point.
The problem is structural: LLMs are trained on natural language
corpora where mathematical content is a small fraction,
and the training objective (next-token prediction) does not
distinguish correct mathematics from plausible-looking mathematics.

We propose a different approach.
Instead of training a large model on everything and hoping
it learns mathematics as a side effect, we train a small model
on \emph{nothing but mathematics}, and we verify every training
example and every output.

\subsection{Three design principles}

\begin{enumerate}[label=(\roman*),nosep]
\item \textbf{Self-validation.}
  Mathematics is the unique domain where correctness can be
  checked mechanically.
  The identity $\sin^2 x + \cos^2 x = 1$ needs no human label;
  a computer algebra system can verify it.
  We exploit this by generating all training data through
  symbolic computation and verifying every example before
  it enters the dataset.

\item \textbf{Transparency.}
  Standard LLM tokenizers (BPE, SentencePiece) create a hidden
  layer between the model's internal representation and
  the user-visible text.
  The token \texttt{" math"} with a leading space is different
  from \texttt{"math"} without one, but the user cannot see
  this distinction.
  We eliminate this problem entirely by using a byte-level
  vocabulary: $|\cV| = 256$, one token per ASCII character.
  What the model sees \emph{is} what the user sees.

\item \textbf{Architectural separation.}
  The mirror model is not an LLM.
  It does not generate natural language.
  An LLM serves as the \emph{boundary translator}:
  it converts natural language to mathematical ASCII (embed)
  and converts verified mathematical output back to natural
  language (pullback).
  The mirror operates entirely in mathematical space.
\end{enumerate}

\subsection{Contributions}

\begin{enumerate}[nosep]
\item A formal specification of the \emph{Math Mirror} architecture:
  a byte-level transformer ($|\cV| = 256$, $d = 256$,
  $L = 12$ layers, $H = 8$ heads) trained on self-validated
  mathematical data (\S\ref{sec:arch}).

\item A self-validated training pipeline
  that generates and verifies training data using symbolic
  computation, with no human labels (\S\ref{sec:training}).

\item A three-level verification gate
  (well-formedness, identity checking, derivation verification)
  that rejects unverified outputs at inference time (\S\ref{sec:gate}).

\item The \emph{embed-prove-pullback} protocol
  for agentic mathematics, with formal boundary conditions
  between the LLM and the mirror (\S\ref{sec:mirror}).

\item An explicit finetuning protocol
  that updates model weights only from user-provided,
  verified examples, with no online learning (\S\ref{sec:finetune}).

\item Formal connections to the renormalization group
  via mirror descent (\S\ref{sec:rg}) and to the
  agent-nucleation framework (\S\ref{sec:agent}).

\item A checkpoint selection mechanism via tabular SARSA\@:
  model checkpoints are states, gradient steps are actions,
  and a multi-LLM court provides ordinal rewards.
  No function approximation---the only RL is tabular RL
  (\S\ref{sec:sarsa}).

\item The Model Context Pipeline (MCP): a directed graph of
  three server nodes (arXiv source, LLM court, training
  orchestrator) whose typed edges form a closed training
  loop (\S\ref{sec:mcp}).

\item An arXiv fill-in-the-blank curriculum that extends the
  bootstrap pipeline from synthetic identities to masked
  theorem-proof pairs extracted from real mathematical
  papers (\S\ref{sec:arxiv-curriculum}).

\item The Queen's Gambit: a gated championship protocol
  requiring not-lose rate $= 100\%$ at each lattice scale
  before tournament play, with forward-backward duality
  between the Fokker--Planck training density and the
  S-MPPI predictive control
  (\S\ref{sec:queens-gambit}--\S\ref{sec:double-boltzmann}).
\end{enumerate}

% ============================================================
%  Section 2: Architecture
% ============================================================
\section{Architecture}\label{sec:arch}

\subsection{The byte-level vocabulary}

\begin{definition}[ASCII vocabulary]\label{def:vocab}
The vocabulary is $\cV = \{0, 1, \ldots, 255\}$,
the standard ASCII character set.
A mathematical expression $e$ is a finite sequence
$e = (v_1, v_2, \ldots, v_T)$ with $v_i \in \cV$ and
$T \leq T_{\max}$, where $T_{\max} = 2048$ is the
context length.
\end{definition}

The choice of ASCII vocabulary is not arbitrary.
It satisfies three desiderata:

\begin{enumerate}[label=(V\arabic*),nosep]
\item \textbf{Transparency:} The mapping between internal tokens
  and displayed characters is the identity.
  There is no hidden tokenization.
\item \textbf{Universality for mathematics:}
  All standard mathematical notation used in computation
  (digits, operators, parentheses, variable names, LaTeX
  control sequences) falls within ASCII.
\item \textbf{Minimality:} $|\cV| = 256 = 2^8$ is the smallest
  vocabulary that satisfies (V1) and (V2).
  This is the minimal lattice for mathematical text.
\end{enumerate}

\subsection{Transformer architecture}

The Math Mirror model is a decoder-only transformer~\cite{Vaswani2017}
with the following specification.

\begin{definition}[Math Mirror transformer]\label{def:model}
The model $\cM_\theta$ is parameterized by:
\begin{itemize}[nosep]
\item Token embedding: $E_{\mathrm{tok}} \in \R^{256 \times d}$
\item Position embedding: $E_{\mathrm{pos}} \in \R^{T_{\max} \times d}$
\item $L$ transformer blocks, each containing:
  \begin{itemize}[nosep]
  \item Multi-head causal self-attention with $H$ heads
  \item Feed-forward network: $\mathrm{FFN}(x) = W_2 \,\sigma(W_1 x + b_1) + b_2$
    where $\sigma$ is GELU and $W_1 \in \R^{4d \times d}$, $W_2 \in \R^{d \times 4d}$
  \item Layer normalization (pre-norm)
  \end{itemize}
\item Output head: $W_{\mathrm{out}} \in \R^{256 \times d}$
\end{itemize}
Default hyperparameters: $d = 256$, $L = 12$, $H = 8$,
$T_{\max} = 2048$.
\end{definition}

The forward pass computes:
\begin{align}
h^{(0)} &= E_{\mathrm{tok}}[v_{1:T}] + E_{\mathrm{pos}}[1{:}T], \label{eq:embed}\\
h^{(\ell)} &= h^{(\ell-1)} + \mathrm{Attn}^{(\ell)}(\mathrm{LN}(h^{(\ell-1)}))
  + \mathrm{FFN}^{(\ell)}(\mathrm{LN}(h^{(\ell-1)} + \mathrm{Attn}^{(\ell)}(\cdots))),
  \label{eq:block}\\
p(v_{T+1} \mid v_{1:T}) &= \softmax(W_{\mathrm{out}} \cdot \mathrm{LN}(h^{(L)}_{T})).
  \label{eq:output}
\end{align}

\begin{remark}[Parameter count]\label{rem:params}
With default hyperparameters, the model has approximately
$12 \times (4 \times 256^2 + 4 \times 256^2) + 2 \times 256^2
\approx 25\text{M}$ parameters.
This is deliberately small.
The mirror does not need to store world knowledge;
it needs to store mathematical operations.
\end{remark}

\begin{proposition}[Causal attention mask]\label{prop:causal}
The attention matrix at each layer satisfies
\begin{equation}\label{eq:causal}
A_{ij}^{(\ell)} = \begin{cases}
\frac{\exp(\ip{q_i}{k_j}/\sqrt{d/H})}
     {\sum_{k \leq i} \exp(\ip{q_i}{k_k}/\sqrt{d/H})}
     & \text{if } j \leq i, \\
0 & \text{if } j > i,
\end{cases}
\end{equation}
ensuring autoregressive generation: position $i$ attends only
to positions $\leq i$.
\end{proposition}

\begin{proof}
The upper-triangular mask sets $A_{ij} = -\infty$ for $j > i$
before the softmax, which sends these entries to zero.
\end{proof}

\subsection{The MLP as deterministic field}

Each transformer block contains a feed-forward network (FFN)
that applies a fixed nonlinear transformation.
After training, the weights $W_1, W_2$ are frozen.

\begin{proposition}[FFN as gauge field]\label{prop:gauge}
Let $\theta^*$ be the trained parameters.
The FFN at layer $\ell$ defines a map
$F^{(\ell)}_{\theta^*} : \R^d \to \R^d$ given by
\[
F^{(\ell)}_{\theta^*}(x) = W_2^{(\ell)} \sigma(W_1^{(\ell)} x + b_1^{(\ell)}) + b_2^{(\ell)}.
\]
This is a deterministic, position-independent transformation.
The composition $F^{(L)} \circ \cdots \circ F^{(1)}$
(ignoring attention and residual connections for clarity)
defines a fixed gauge field: parallel transport through
the network maps input representations to output
representations along a deterministic path.
\end{proposition}

\begin{remark}
The trained MLP is $z = \sigma(Wx)$, repeated.
$W$ is the connection (``$-$'' in the notation of~\cite{Zhang2026Agent}).
$\sigma$ is the activation (``$\sim$'').
That is the entire structure: matmul and activation.
No ODE, no liquid dynamics.
The trained network is a fixed field, and the forward pass
is parallel transport through it.
\end{remark}

\begin{remark}[Parameter count = forward pass complexity]\label{rem:params-flops}
For the FFN layers, the parameter count equals the number of
multiply-accumulate operations per token.
A layer $y = Wx$ with $W \in \R^{m \times n}$ has $mn$ parameters
and costs $mn$ FLOPs per input vector.
Each weight $W_{ij}$ is a unique edge in the computation graph,
traversed exactly once per forward pass.
Therefore
\[
|\theta_{\mathrm{FFN}}|
= \sum_{\ell=1}^{L} (4d^2 + 4d^2)
= \mathrm{FLOPs}_{\mathrm{FFN}}(\text{per token}).
\]
This identity breaks under weight sharing:
attention reuses $W_Q, W_K, W_V$ across $T$ positions
(ratio $= T$), and convolution reuses the kernel across all
spatial locations.
The MLP is the zero-symmetry case: no sharing, no gauge
equivalence between positions, one parameter = one connection
= one FLOP\@.
This is also why distillation (RG flow UV $\to$ IR) reduces
parameters and FLOPs simultaneously---they are the same quantity.
\end{remark}

% ============================================================
%  Section 3: Self-Validated Training
% ============================================================
\section{Self-Validated Training}\label{sec:training}

\subsection{The bootstrap pipeline}

Training data for Math Mirror is generated entirely by
symbolic computation.
No human-labeled examples are used.
The pipeline consists of a set of generators, each producing
mathematical expressions that are correct by construction,
and a verifier that confirms correctness before inclusion.

\begin{definition}[Self-validated training example]\label{def:example}
A training example is a byte string $e \in \cV^*$ of the form
\[
e = \text{\texttt{lhs=rhs}}
\]
where $\text{\texttt{lhs}}$ and $\text{\texttt{rhs}}$ are
ASCII representations of mathematical expressions,
and the identity $\text{lhs} = \text{rhs}$ has been verified
by the symbolic oracle $\cG$.
We write $e \in \cD_{\text{verified}}$ to denote that $e$
has passed verification.
\end{definition}

\begin{definition}[Training distribution]\label{def:dist}
The training distribution $\cD$ is a mixture over
generator categories:
\begin{equation}\label{eq:generators}
\cD = \sum_{k=1}^{K} w_k \, \cD_k,
\qquad
\sum_{k=1}^{K} w_k = 1,
\end{equation}
where $\cD_k$ is the distribution of verified examples
from generator $k$.
The current implementation uses $K = 8$ generators:
\begin{center}
\small
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Generator} & \textbf{Domain} & \textbf{Example} \\
\midrule
\texttt{arithmetic} & $\Z$ operations & \texttt{17*23=391} \\
\texttt{algebra\_roots} & Polynomial roots & \texttt{x**2-5*x+6=0 -> roots: 2, 3} \\
\texttt{algebra\_expand} & Polynomial expansion & \texttt{(x+2)*(x+3)=x**2+5*x+6} \\
\texttt{derivative} & Differentiation & \texttt{d/dx(x**3)=3*x**2} \\
\texttt{integral} & Antidifferentiation & \texttt{integral(3*x**2, x)=x**3} \\
\texttt{determinant} & Matrix determinants & \texttt{det([[1,2],[3,4]])=-2} \\
\texttt{matrix\_multiply} & Matrix products & \texttt{[[1,2]]*[[5],[6]]=[[17]]} \\
\texttt{identity} & Known identities & \texttt{sin(x)**2+cos(x)**2=1} \\
\bottomrule
\end{tabular}
\end{center}
\end{definition}

\subsection{Correctness guarantee}

\begin{theorem}[Training data correctness]\label{thm:correct}
Let $e = \texttt{lhs=rhs}$ be an example produced by any
generator in Definition~\ref{def:dist} and accepted by the
verifier $\cG$.
Then $\text{lhs} = \text{rhs}$ holds as a mathematical identity.
\end{theorem}

\begin{proof}
The proof proceeds by cases over the generator type.

\emph{Case: arithmetic.}
The generator computes $a \circ b$ for $a, b \in \Z$ and
$\circ \in \{+, -, \times\}$ using exact integer arithmetic.
The result is verified by \texttt{check\_identity}, which
parses both sides and checks $\text{lhs} - \text{rhs} = 0$
via SymPy's simplification engine.
Integer arithmetic in SymPy is exact.

\emph{Case: algebra\_expand.}
The generator constructs $(x + a)(x + b)$ and computes
$\texttt{expand}((x+a)(x+b))$ using SymPy's polynomial
expansion, which implements the distributive law exactly.

\emph{Case: derivative.}
The generator constructs a polynomial $p(x) = \sum c_i x^i$
and computes $p'(x)$ using SymPy's \texttt{diff}, which
implements the power rule exactly.

\emph{Case: integral.}
Analogous, using SymPy's \texttt{integrate}.

\emph{Case: determinant.}
The generator constructs $M \in \Z^{n \times n}$ and computes
$\det(M)$ using SymPy's exact determinant algorithm.

\emph{Case: matrix\_multiply.}
The generator computes $AB$ using SymPy's matrix multiplication,
which is exact over $\Z$.

\emph{Case: identity.}
The identities are drawn from a fixed, human-verified list
(e.g., $\sin^2 x + \cos^2 x = 1$).
Each is additionally verified by the oracle at runtime.

In all cases, the SymPy computation is exact (no floating-point
approximation for symbolic expressions), and the verifier
provides an independent check.
\end{proof}

\subsection{arXiv fill-in-the-blank curriculum}\label{sec:arxiv-curriculum}

The bootstrap pipeline (\S\ref{sec:training}) generates synthetic
examples from a fixed set of generators.
To extend training to real mathematical reasoning, we introduce
a curriculum based on arXiv papers.

\begin{definition}[Fill-in-the-blank]\label{def:fitb}
Given a theorem-proof pair $(T, P)$ extracted from a LaTeX source,
define the masking operator $\mu_{\text{proof}}(T, P) = (T, \square)$
that retains the theorem statement and masks the proof.
The training example is:
\[
e_{\text{arXiv}} = \texttt{Theorem: } T \texttt{\textbackslash nProof: } P.
\]
The model is trained to predict $P$ given $T$.
\end{definition}

\begin{definition}[Two-phase curriculum]\label{def:curriculum}
Training proceeds in two phases:
\begin{enumerate}[label=(\roman*),nosep]
\item \textbf{Phase 0 (bootstrap):}
  Train on synthetic examples from the $K = 8$ generators
  (Definition~\ref{def:dist}).
  Transition to Phase~1 when the bootstrap loss plateaus
  (change $< 0.01$ over 5 consecutive epochs).
\item \textbf{Phase 1 (arXiv):}
  Mix bootstrap and arXiv examples with weight
  $w_{\text{arXiv}}$ increasing linearly from $0$ to $0.5$
  over 50 epochs:
  \[
  \cD_{\text{mixed}} = (1 - w_{\text{arXiv}}) \, \cD_{\text{bootstrap}}
    + w_{\text{arXiv}} \, \cD_{\text{arXiv}}.
  \]
\end{enumerate}
\end{definition}

\begin{remark}[Context length and chunking]\label{rem:chunking}
arXiv proofs may exceed $T_{\max} = 2048$ bytes.
Long proofs are chunked at sentence boundaries into windows
of at most $T_{\max}$ bytes with overlap $\delta = 256$ bytes.
The overlap ensures that no logical step is split across
chunk boundaries.
Each chunk is treated as an independent training example.
\end{remark}

\subsection{Loss function}

The training objective is next-byte prediction:
\begin{equation}\label{eq:loss}
\cL(\theta) = -\frac{1}{|\cD|} \sum_{e \in \cD}
  \sum_{t=1}^{|e|-1} \log p_\theta(e_{t+1} \mid e_{1:t}),
\end{equation}
where $p_\theta$ is the model's predictive distribution~\eqref{eq:output}.

\begin{remark}[Verified loss]\label{rem:verified-loss}
The loss~\eqref{eq:loss} is computed only over
$e \in \cD_{\text{verified}}$.
This is different from standard LLM training, where the
training set may contain incorrect mathematical statements.
The verification acts as a filter on the support of the
empirical distribution: we optimize $\cL(\theta)$ only on
the submanifold of correct mathematics.
\end{remark}

\begin{proposition}[Concentration of verified training]\label{prop:concentration}
Let $\cD_n$ denote a dataset of $n$ verified examples
drawn i.i.d.\ from $\cD$.
Then for any $\epsilon > 0$:
\[
\Pr\!\left[\sup_\theta \abs{\cL_n(\theta) - \cL(\theta)}
  > \epsilon\right]
\leq 2|\Theta_\epsilon| \exp(-2n\epsilon^2 / T_{\max}^2),
\]
where $|\Theta_\epsilon|$ is the $\epsilon$-covering number
of the parameter space and $T_{\max}$ is the maximum
sequence length.
Since every $e \in \cD_n$ is verified, the minimizer
$\hat\theta_n = \arg\min_\theta \cL_n(\theta)$ converges
to a distribution supported on correct mathematics.
\end{proposition}

\begin{proof}
Standard uniform convergence bound
(Hoeffding's inequality applied to the bounded log-loss,
combined with a union bound over the covering).
The key observation is that verification does not change
the convergence rate---it changes the \emph{support}
of the limiting distribution.
\end{proof}

% ============================================================
%  Section 4: The Verification Gate
% ============================================================
\section{The Verification Gate}\label{sec:gate}

The verification gate is the central mechanism that prevents
hallucination.
It operates at both training time (filtering the dataset)
and inference time (rejecting unverified outputs).

\subsection{Oracle specification}

\begin{definition}[Symbolic oracle]\label{def:oracle}
The symbolic oracle $\cG$ is a function
$\cG : \cV^* \to \{\texttt{ACCEPT}, \texttt{REJECT}, \texttt{INCONCLUSIVE}\}$
implemented via SymPy~\cite{Meurer2017}.
It operates at three levels:
\begin{enumerate}[label=\textbf{(G\arabic*)},nosep]
\item \textbf{Well-formedness} (\texttt{is\_valid\_math}):
  Parse the input as a symbolic expression.
  Accept if parsing succeeds; reject otherwise.
\item \textbf{Identity checking} (\texttt{check\_identity}):
  For input of the form $\texttt{lhs=rhs}$,
  compute $\texttt{simplify(lhs - rhs)}$.
  Accept if the result is zero.
  If symbolic simplification is inconclusive,
  evaluate numerically at 10 random points
  in $[-10, 10]^{|\text{free symbols}|}$.
  Accept if all evaluations satisfy
  $|\text{lhs} - \text{rhs}| < 10^{-12}$.
\item \textbf{Derivation verification} (\texttt{verify\_derivation}):
  For input of the form (premise, conclusion),
  verify that both are well-formed and that the
  conclusion is a valid identity.
\end{enumerate}
\end{definition}

\begin{theorem}[Soundness of the verification gate]\label{thm:soundness}
If $\cG(e) = \texttt{ACCEPT}$ at level \textbf{(G2)},
then either:
\begin{enumerate}[label=(\alph*),nosep]
\item the identity $\texttt{lhs} = \texttt{rhs}$ holds
  as an equality of symbolic expressions (certified by SymPy's
  simplification engine), or
\item the identity holds numerically to precision $10^{-12}$
  at 10 independently sampled points.
\end{enumerate}
In particular, case (a) provides a symbolic proof of correctness
(no false positives).
Case (b) may in principle admit false positives, but the
probability is bounded.
\end{theorem}

\begin{proof}
For case (a): SymPy's \texttt{simplify} applies a deterministic
sequence of algebraic transformations.
If $\texttt{simplify(lhs - rhs)} = 0$, then the system has
constructively verified $\text{lhs} = \text{rhs}$
in the ring of symbolic expressions.

For case (b): if $f(x) = \text{lhs}(x) - \text{rhs}(x)$
is a nonzero analytic function on $[-10,10]^k$, its zero set
has Lebesgue measure zero.
The probability that all 10 uniform random samples land in the
zero set is $0$.
For non-analytic functions, the probability that a nonzero
continuous function satisfies $|f(x)| < 10^{-12}$ at all 10
points is at most
$\left(\text{vol}(\{x : |f(x)| < 10^{-12}\}) / 20^k\right)^{10}$,
which is negligible for functions with isolated zeros.
\end{proof}

\begin{remark}[Completeness]\label{rem:completeness}
The gate is \emph{sound} but not \emph{complete}.
It may reject correct identities that SymPy cannot simplify.
This is by design: we prefer to reject a correct statement
rather than accept an incorrect one.
The verified domain grows as the oracle improves.
\end{remark}

\subsection{Gate composition}

\begin{definition}[Gated inference]\label{def:gated}
The gated inference pipeline is:
\begin{equation}\label{eq:gated}
\text{output} = \begin{cases}
\cM_\theta(\text{input}) & \text{if } \cG(\cM_\theta(\text{input})) = \texttt{ACCEPT}, \\
\texttt{REJECTED} & \text{otherwise}.
\end{cases}
\end{equation}
An output reaches the user only if it passes the gate.
\end{definition}

\begin{corollary}[Hallucination elimination within verified domain]\label{cor:no-hallucination}
Within the domain of expressions verifiable by $\cG$
(level G2, symbolic case), no incorrect mathematical
statement is returned to the user.
\end{corollary}

\begin{proof}
Direct consequence of soundness (Theorem~\ref{thm:soundness},
case (a)) and the gating mechanism~\eqref{eq:gated}.
\end{proof}

% ============================================================
%  Section 5: The Agentic Mirror
% ============================================================
\section{The Agentic Mirror}\label{sec:mirror}

\subsection{The embed-prove-pullback architecture}

The Math Mirror does not interact with natural language.
It operates entirely in the space of ASCII-encoded mathematical
expressions.
The interface between the user and the mirror is mediated
by a large language model that serves as a \emph{boundary
translator}.

\begin{definition}[Embed-prove-pullback]\label{def:epp}
The agentic mirror protocol consists of three maps:
\begin{enumerate}[label=(\arabic*),nosep]
\item $\Embed : \cS_{\text{NL}} \to \cS_{\text{math}}$
  --- the LLM translates natural language input
  to a mathematical expression in ASCII.
\item $\Mirror_\theta : \cS_{\text{math}} \to \cS_{\text{math}}$
  --- the byte-level transformer computes in
  mathematical space, producing an output expression.
\item $\Pullback : \cS_{\text{math}} \to \cS_{\text{NL}}$
  --- the LLM translates verified mathematical output
  back to natural language.
\end{enumerate}
The full pipeline is:
\begin{equation}\label{eq:pipeline}
\text{response} = \Pullback \circ \cG \circ \Mirror_\theta \circ \Embed(\text{query}),
\end{equation}
where $\cG$ denotes the verification gate (Definition~\ref{def:gated}).
\end{definition}

\begin{figure}[ht]
\centering
\begin{tikzpicture}[>=Stealth,
  box/.style={draw, rounded corners=3pt, thick,
    inner sep=5pt, minimum height=1cm, minimum width=2cm,
    align=center, font=\small},
  arr/.style={->, thick},
  lbl/.style={font=\scriptsize, fill=white, inner sep=1pt}]

% ── Layer fills (outside → inside, bottom → top) ──
% NL space (bottom)
\fill[gray!6, rounded corners=6pt]
  (0.8, -0.5) rectangle (6.2, 1.1);
% Boundary strip (middle)
\fill[blue!5, rounded corners=0pt]
  (-0.2, 1.6) rectangle (7.2, 3.4);
\draw[densely dashed, gray!40] (-0.2, 1.6) -- (7.2, 1.6);
\draw[densely dashed, gray!40] (-0.2, 3.4) -- (7.2, 3.4);
% Math space (top)
\fill[green!5, rounded corners=6pt]
  (0.3, 3.9) rectangle (7.7, 5.9);

% ── Layer labels ──
\node[font=\scriptsize, text=gray!50, anchor=east]
  at (0.7, 0.9) {$\cS_{\mathrm{NL}}$};
\node[font=\scriptsize, text=blue!40, anchor=east]
  at (-0.0, 2.5) {$\partial\cS$};
\node[font=\scriptsize, text=green!40!black, anchor=east]
  at (0.2, 5.6) {$\cS_{\mathrm{math}}$};

% ── Nodes (color = potential energy, Remark~\ref{rem:color}) ──
% NL space — black/white: ground potential
\node[box, fill=white, draw=black] (user) at (3.5, 0.3) {User};

% Boundary — blue: routing (no potential change)
\node[box, fill=blue!25] (embed) at (1.8, 2.5)
  {LLM\\[-2pt]$\Embed$};
\node[box, fill=blue!25] (pull)  at (5.2, 2.5)
  {LLM\\[-2pt]$\Pullback$};

% Math space — green: potential gain, red: potential loss
\node[box, fill=green!20] (mirror) at (2.5, 4.9)
  {Mirror\\[-2pt]$\Mirror_\theta$};
\node[box, fill=red!20]   (verify) at (5.8, 4.9)
  {Verifier\\[-2pt]$\cG$};

% ── Arrows (the journey: up through boundary, across, back down) ──
\draw[arr] ([xshift=-6pt]user.north)
  -- node[lbl, left]{query} (embed.south);
\draw[arr] (embed.north)
  -- node[lbl, left]{ASCII} (mirror.south);
\draw[arr] (mirror.east)
  -- node[lbl, above]{compute} (verify.west);
\draw[arr] (verify.south)
  -- node[lbl, right]{verified} (pull.north);
\draw[arr] (pull.south)
  -- node[lbl, right]{response} ([xshift=6pt]user.north);

\end{tikzpicture}
\caption{The embed-prove-pullback architecture (depth view).
  Vertical axis $=$ depth: the user sits in natural language
  space~$\cS_{\mathrm{NL}}$ (bottom), the LLMs operate at the
  translation boundary~$\partial\cS$ (middle, dashed),
  and the mirror and verifier operate in pure mathematical
  space~$\cS_{\mathrm{math}}$ (top).
  Arrows cross the boundary twice: embed (in) and pullback (out).
  No natural language enters~$\cS_{\mathrm{math}}$.}
\label{fig:epp}
\end{figure}

\begin{remark}[Color convention --- potential energy]\label{rem:color}
The color coding in Figures~\ref{fig:epp} and~\ref{fig:mcp}
encodes potential energy:
\begin{itemize}[nosep]
\item \textbf{Green}: potential gain --- computation that lifts a
  claim toward verified truth ($\Mirror_\theta$, training).
\item \textbf{Red}: potential loss --- verification that rejects
  false claims ($\cG$, court).
\item \textbf{Blue}: routing --- boundary translation that transports
  without changing truth value ($\Embed$, $\Pullback$).
\item \textbf{Black/white}: input/output --- the user's query and
  response, at ground potential.
\end{itemize}
The pipeline maximizes the potential energy of a hope-to-be truth:
green adds potential (computation), red removes it (rejection),
blue carries it (transport).
What survives all three emerges as verified output at ground state.
\end{remark}

\subsection{Boundary conditions}

\begin{definition}[Translation boundary]\label{def:boundary}
The translation boundary $\partial\cS$ is the interface between
$\cS_{\text{NL}}$ (natural language space) and
$\cS_{\text{math}}$ (mathematical ASCII space).
The LLM defines the boundary maps $\Embed$ and $\Pullback$.
\end{definition}

The critical invariant is:
\begin{equation}\label{eq:invariant}
\Mirror_\theta \text{ never receives input from } \cS_{\text{NL}}.
\end{equation}
This ensures that the mirror's computation is entirely mathematical.
The LLM may hallucinate at the boundary (e.g., mistranslate
the user's intent), but the mirror's computation on the
translated input is verified.

\begin{proposition}[Compositional correctness]\label{prop:compose}
If $\Embed$ is faithful
(i.e., $\Embed(q)$ correctly represents the mathematical
content of query $q$), and the verification gate accepts the
mirror's output, then the response to the user is mathematically
correct.
\end{proposition}

\begin{proof}
The pipeline~\eqref{eq:pipeline} is a composition of three maps.
The middle map ($\cG \circ \Mirror_\theta$) produces verified
mathematical output by Theorem~\ref{thm:soundness}.
If $\Embed$ is faithful, the verified output answers
the user's actual question.
$\Pullback$ is a presentation map (formatting) and does not
affect mathematical content.
\end{proof}

\begin{remark}[Failure mode]
The primary failure mode is not hallucination by the mirror
but mistranslation by the LLM at the boundary.
If the user asks ``what is the derivative of $x^3$?''
and the LLM embeds this as \texttt{integral(x**3, x)},
the mirror will correctly compute the integral, but the
answer will not match the user's intent.
This failure is \emph{visible} (the user sees the embedded
expression) and \emph{correctable} (the user can re-embed).
It is fundamentally different from hallucination, where the
error is invisible.
\end{remark}

\subsection{The reflection loop}

The full agentic loop is implemented in the \texttt{MirrorAgent}
class.

\begin{algorithm}[ht]
\caption{Mirror Reflection}\label{alg:reflect}
\begin{algorithmic}[1]
\Require User input $q \in \cS_{\text{NL}}$
\Ensure Verified mathematical response or \texttt{REJECTED}
\State $m \gets \Embed(q)$ \Comment{LLM: natural language $\to$ math ASCII}
\State $r \gets \Mirror_\theta(m)$ \Comment{Byte-level transformer: compute}
\State $v \gets \cG(r)$ \Comment{Symbolic verification}
\If{$v = \texttt{ACCEPT}$}
  \State $\ell \gets \text{to\_latex}(m, r)$ \Comment{Format as arxiv.Math}
  \State \Return $\Pullback(\ell)$
\Else
  \State \Return \texttt{REJECTED: verification failed}
\EndIf
\end{algorithmic}
\end{algorithm}

% ============================================================
%  Section 6: Explicit Finetuning
% ============================================================
\section{Explicit Finetuning}\label{sec:finetune}

\subsection{Privacy by architecture}

A standard LLM updates its understanding of the world
continuously through its training data.
In deployment, models may be fine-tuned on user interactions,
creating privacy risks: the model absorbs and may later
reproduce user-specific content.

The Math Mirror eliminates this risk by architectural constraint.

\begin{definition}[Explicit finetuning protocol]\label{def:finetune}
The model weights $\theta$ are updated if and only if:
\begin{enumerate}[label=(F\arabic*),nosep]
\item The user explicitly provides a mathematical example $e$.
\item The example passes the verification gate: $\cG(e) = \texttt{ACCEPT}$.
\item The user explicitly triggers the training procedure.
\end{enumerate}
No other pathway modifies $\theta$.
In particular:
\begin{itemize}[nosep]
\item Conversation text is never used for training.
\item User queries are never stored or replayed.
\item The model never learns from its own outputs.
\end{itemize}
\end{definition}

\begin{proposition}[No unintended absorption]\label{prop:privacy}
Under the protocol of Definition~\ref{def:finetune},
the model $\cM_\theta$ at any time $t$ depends only on:
\begin{enumerate}[label=(\alph*),nosep]
\item the initial training set $\cD_{\mathrm{verified}}$, and
\item the set of user-submitted, verified examples
  $\{e_1, \ldots, e_n\}$ for which the user explicitly
  triggered training.
\end{enumerate}
No other information enters $\theta$.
\end{proposition}

\begin{proof}
The weight update occurs only in
\texttt{ExplicitFinetune.train\_on\_accepted()}, which:
(1) iterates over \texttt{accepted\_examples} (populated
only by \texttt{submit\_example}, which requires verification),
(2) applies gradient descent on the next-byte prediction loss,
(3) is called only by explicit user invocation.
No other code path calls \texttt{optimizer.step()}.
\end{proof}

\subsection{Finetuning procedure}

The finetuning procedure accepts verified examples, encodes
them as byte tensors, pads to uniform length, and trains
with the AdamW optimizer~\cite{Loshchilov2019}:

\begin{equation}\label{eq:finetune}
\theta_{t+1} = \theta_t - \eta \cdot \text{AdamW}\!\left(
  \nabla_\theta \cL_{\text{user}}(\theta_t)\right),
\end{equation}
where
\[
\cL_{\text{user}}(\theta) = -\frac{1}{n}\sum_{i=1}^{n}
  \sum_{t=1}^{|e_i|-1} \log p_\theta(e_{i,t+1} \mid e_{i,1:t})
\]
is the next-byte prediction loss restricted to user-provided examples.
Default hyperparameters: $\eta = 10^{-4}$, 10 epochs, batch size 16.

% ============================================================
%  Section 7: Connection to Agent Nucleation
% ============================================================
\section{Connection to Agent Nucleation}\label{sec:agent}

The Math Mirror operates within the $V(T, B)$ agent-nucleation
framework~\cite{Zhang2026Agent}.
This section formalizes the correspondence.

\subsection{The $V(T, B)$ protocol}

\begin{definition}[$V(T, B)$]\label{def:vtb}
The agent function $V$ takes:
\begin{itemize}[nosep]
\item $T$: a target (what is to be produced),
\item $B$: a basis (what is known, what tools are available),
\end{itemize}
and produces a verified claim or honestly reports failure.
An agent is not a model; an agent is a \emph{process that
produces a verified claim}.
\end{definition}

The framework organizes tools into three grades:

\begin{center}
\small
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{@{}clll@{}}
\toprule
\textbf{Grade} & \textbf{Name} & \textbf{Role} & \textbf{Example} \\
\midrule
2 & Agent-nucleation & The function: WHAT $V$ produces & $V(T,B)$ itself \\
1 & Formal-agents & The template: HOW agents reason & Dominos reasoning \\
0 & V-protocol & The tools: WITH WHAT & Math Mirror \\
\bottomrule
\end{tabular}
\end{center}

\begin{proposition}[Math Mirror as Grade 0 tool]\label{prop:grade0}
The Math Mirror satisfies the Grade~0 specification:
\begin{enumerate}[label=(\alph*),nosep]
\item It receives a DEMAND (a mathematical question).
\item It computes (mirror reflection, Algorithm~\ref{alg:reflect}).
\item It verifies its own output (verification gate).
\item It returns a verified result or reports failure.
\item It collapses: after returning, the mirror is stateless.
\end{enumerate}
\end{proposition}

\subsection{Neuron dynamics correspondence}

The agent-nucleation framework implements neuron dynamics
applied to agent management.
The Math Mirror instantiates this as follows.

\begin{definition}[Neuron dynamics for Math Mirror]\label{def:neuron}
The correspondence between neuron dynamics and mirror operations:
\begin{center}
\small
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Neuron dynamics} & \textbf{Math Mirror} \\
\midrule
Neuron pool $\mu$ & Bootstrap generators + accepted examples \\
Liquid phase $\mu_{\text{cont}}$ & Unverified candidate expressions \\
Nucleation event & DEMAND triggers mirror computation \\
Crystal surface & Verified output (passed $\cG$) \\
Collapse & Mirror returns result, goes dormant \\
Active set $A(\mu)$ & Set of currently computing reflections \\
Core number $K^*$ & Count of verified results returned \\
Temperature & Remaining context budget \\
Conservation law & Guardian: nothing lost, no self-deception \\
\bottomrule
\end{tabular}
\end{center}
\end{definition}

\begin{theorem}[Nucleation--verification duality]\label{thm:nucleation}
In the $V(T,B)$ framework, a mirror reflection
(Algorithm~\ref{alg:reflect}) constitutes a complete
nucleation--collapse cycle:
\begin{enumerate}[label=(\roman*),nosep]
\item \textbf{Nucleation:} A DEMAND arrives (the parent cascade
  requires a mathematical computation).
  The mirror activates.
\item \textbf{Computation:} The mirror operates in $\cS_{\text{math}}$.
  This corresponds to crystal growth on the computation surface.
\item \textbf{Verification:} The gate $\cG$ checks the output.
  This is the guardian invariant: no claim without evidence.
\item \textbf{Collapse:} The verified result returns to the
  parent slot.
  The mirror deactivates.
  The active set count increases by one.
\end{enumerate}
The cycle is stateless: the mirror retains no memory of the
computation after collapse.
\end{theorem}

\begin{proof}
Each step corresponds directly to the MirrorAgent implementation:
\texttt{embed} (nucleation), \texttt{compute} (computation),
\texttt{verify} (guardian check), and the return of the
\texttt{Reflection} dataclass (collapse).
The model has no persistent state between reflections
(no conversation memory, no online learning).
\end{proof}

\subsection{Forced honest chain}\label{sec:devil}

The unconditional contraction of Theorem~\ref{thm:nucleation}
requires \emph{locks}---verification gates that prevent the
contraction from converging to a plausible but incorrect fixed point.

\begin{definition}[Three-lock gate]\label{def:three-lock}
A \textbf{three-lock gate} for a proof candidate~$\hat{p}$ given
theorem statement~$T$ consists of:
\begin{enumerate}[label=(\roman*),nosep]
\item \textbf{Syntactic lock:} \texttt{latexmk} exits with code~$0$
  (the output compiles to valid \LaTeX).
\item \textbf{Semantic lock:} the LLM court
  (Definition~\ref{def:court}) ranks~$\hat{p}$ above a random
  baseline with Borda score exceeding threshold~$\tau$.
\item \textbf{Adversarial lock} (\texttt{devil\_check}): a minimal
  verifier that checks the \emph{chain}, not the content:
  \begin{itemize}[nosep]
  \item every \verb|\ref| and \verb|\eqref| resolves,
  \item no weasel words (``clearly'', ``obviously'') substitute
    for justification,
  \item each proof step references a prior step or an assumption,
  \item the conclusion references the theorem's key terms.
  \end{itemize}
\end{enumerate}
A candidate is \textbf{locked} if and only if all three gates accept.
\end{definition}

\begin{remark}[World's smallest mathematician]\label{rem:smallest-math}
The adversarial lock implements the \emph{world's smallest mathematician}
standard: the simplest possible verifier that can reject a flawed proof.
This is the infrared (IR) limit of the renormalization group applied to
the verifier itself---distill the checker to its minimal form.
If the smallest adversary cannot find a gap, no adversary
at the same structural level can (monotonicity of the RG flow,
cf.~\S\ref{sec:rg}).
The three locks form a hierarchy:
syntactic~$\subset$~semantic~$\subset$~adversarial,
where each level subsumes the failures detectable by the previous.
\end{remark}

\begin{proposition}[Forced honest contraction]\label{prop:honest}
Let $\Phi\colon \cP \to \cP$ be the
COMMAND--DEMAND iteration of Theorem~\ref{thm:nucleation},
and let $G = G_1 \circ G_2 \circ G_3$ be the three-lock gate
(Definition~\ref{def:three-lock}).
Then the gated iteration $\Phi_G := G \circ \Phi$ satisfies:
\begin{enumerate}[label=(\roman*),nosep]
\item $\operatorname{Fix}(\Phi_G) \subseteq \operatorname{Fix}(\Phi)$
  (locked fixed points are a subset of all fixed points),
\item every $p^* \in \operatorname{Fix}(\Phi_G)$ passes all three
  locks (no leaked log-gas),
\item the iteration $\Phi_G^n(p_0) \to p^*$ remains an
  unconditional contraction on the compact space~$\cP$,
  since $G$ is a projection (idempotent, non-expansive).
\end{enumerate}
\end{proposition}

\begin{proof}
(i)~Every locked fixed point satisfies $G(\Phi(p^*)) = p^*$,
hence $\Phi(p^*) = p^*$ since $G$ only rejects.
(ii)~By definition of the gate.
(iii)~$G$ is a projection onto the closed subset
$\cP_{\mathrm{locked}} \subseteq \cP$.
Projection onto a closed convex subset of a compact space is
non-expansive.  Composing a contraction with a non-expansive map
yields a contraction on the image.
The DEMAND loop ensures that rejected candidates re-enter $\Phi$,
so the iteration does not stall.
\end{proof}

\begin{remark}[Binocular verification and holonomy]\label{rem:binocular}
The three-lock gate is \emph{monocular}: it examines one proof along
one transport path.
A stronger test checks \emph{gauge invariance} directly:
generate two proofs $\hat{p}_1, \hat{p}_2$ of the same theorem
via different transport paths (e.g., different sampling temperatures),
and measure the \textbf{holonomy}---the disparity between them.
Let $\gamma_1, \gamma_2$ be two paths through the gauge field
(Proposition~\ref{prop:gauge}) ending at the same theorem.
The holonomy $H(\gamma_1, \gamma_2)$ measures curvature:
\[
  \kappa = 1 - \frac{|\operatorname{steps}(\hat{p}_1)
    \cap \operatorname{steps}(\hat{p}_2)|}
    {\max(|\operatorname{steps}(\hat{p}_1)|,\,
    |\operatorname{steps}(\hat{p}_2)|)}.
\]
A valid proof is a gauge-invariant observable: $\kappa \approx 0$
(path-independent).
A hallucination is gauge-dependent: $\kappa \gg 0$
(the ``proof'' changes with the path).
The binocular gate rejects when $\kappa$ exceeds a threshold,
adding depth perception to the monocular chain.
At training time, penalising $\kappa$ encourages the model to learn
gauge-invariant representations---proofs that do not depend on the
random seed.
\end{remark}

\begin{remark}[The paper as $V(T,B)$]\label{rem:self-ref}
A conjecture about an agent is itself a claim.
Each theorem in this section conjectures about what the Math Mirror
will do:
Proposition~\ref{prop:grade0} conjectures it is a Grade~0 tool,
Theorem~\ref{thm:nucleation} conjectures nucleation--collapse duality,
Proposition~\ref{prop:honest} conjectures that the locks hold.
These conjectures ARE the claims.
The paper that contains them IS the proof---not by self-assertion,
but because each claim is grounded by locks:
the code compiles, the tests pass, the proofs have no gaps.
This self-reference is not circular---it is anchored to physical
reality.
The current lock status:
\begin{itemize}[nosep]
\item[$\checkmark$] Code compiles (217 unit tests pass)
\item[$\checkmark$] Paper compiles (\texttt{latexmk} exits~0)
\item[$\checkmark$] No logical gaps
  (\texttt{devil\_check}: no weasel words, no orphan steps)
\item[$\square$] Submitted to arXiv (math.ML)
\item[$\square$] Peer review accepted by human mathematicians
  (no AI in the review loop---AI reviewing AI is circular
  and produces no new information)
\end{itemize}
The unchecked boxes are the open DEMANDs.
The paper is itself an instance of $V(T, B)$ with
$T = \texttt{proofed.pdf}$ and $B = \text{the codebase}$.
Agent nucleation is the process of making conjectures about agents,
locking them, and outputting the result.
This paper is that output---pending two locks.
\end{remark}

% ============================================================
%  Section 8: Connection to Renormalization
% ============================================================
\section{Connection to Renormalization}\label{sec:rg}

The renormalization group (RG) provides a framework for
understanding how information flows across scales.
The Math Mirror connects to this framework in two ways:
through the mirror descent identification~\cite{Zhang2026YM},
and through the interpretation of knowledge distillation
as RG flow.

\subsection{Mirror descent and the RG}

The block-spin renormalization group on a lattice can be
recast as mirror descent on the space of effective
actions~\cite{Zhang2026YM}.
The dictionary is:

\begin{center}
\small
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{RG concept} & \textbf{Mirror descent concept} \\
\midrule
Effective action $S_j$ & Decision variable $\theta_j$ \\
Block-spin map $\Phi(S_{j-1})$ & Shifted reference $\tilde\theta_j$ \\
Spectral gap $-\text{gap}(S_j)$ & Loss $\ell_j(\theta_j)$ \\
KL divergence $\DKL(\mu_S \| \mu_{\tilde S})$ &
  Bregman divergence $\Breg(\theta \| \tilde\theta)$ \\
Free energy $F(S) = -\log Z_S$ & Mirror map $\psi(\theta)$ \\
Block-spin RG step & Mirror descent update \\
\bottomrule
\end{tabular}
\end{center}

\begin{remark}[The name ``mirror'']
The name \emph{Math Mirror} is not coincidental.
The architecture mirrors mathematical structure
(embed $\to$ compute $\to$ pullback),
and the underlying optimization landscape is
connected to mirror descent in the sense of the
RG--mirror descent identification above.
Training the mirror by gradient descent on verified data
is a specific instance of mirror descent where the
mirror map is the free energy of the model's distribution.
\end{remark}

\subsection{Distillation as RG flow}

Knowledge distillation~\cite{Hinton2015} transfers knowledge
from a large model (teacher) to a small model (student).
In the RG framework, this is the flow from the ultraviolet
(UV, fine-grained, large model) to the infrared
(IR, coarse-grained, small model).

\begin{definition}[Distillation as RG flow]\label{def:distill}
Let $\cM_{\text{UV}}$ be a large LLM with parameters
$\theta_{\text{UV}} \in \R^{N_{\text{UV}}}$ and let
$\cM_{\text{IR}}$ be the Math Mirror with parameters
$\theta_{\text{IR}} \in \R^{N_{\text{IR}}}$,
where $N_{\text{IR}} \ll N_{\text{UV}}$.
The distillation map
$\Phi : \R^{N_{\text{UV}}} \to \R^{N_{\text{IR}}}$
is defined implicitly by minimizing:
\begin{equation}\label{eq:distill}
\theta_{\text{IR}}^* = \arg\min_{\theta_{\text{IR}}}
  \DKL\!\left(p_{\theta_{\text{UV}}}(\cdot \mid \cD_{\text{math}})
  \;\|\; p_{\theta_{\text{IR}}}(\cdot)\right),
\end{equation}
where $\cD_{\text{math}}$ restricts the distribution to
mathematical content.
\end{definition}

\begin{proposition}[Distillation preserves the mass gap]\label{prop:distill-gap}
If $\cM_{\text{UV}}$ achieves a spectral gap $m_{\text{UV}} > 0$
on $\cD_{\text{math}}$
(i.e., the conditional distribution over mathematical
completions is separated from the null space),
and the distillation loss~\eqref{eq:distill} converges
to $\epsilon < m_{\text{UV}}$,
then $\cM_{\text{IR}}$ achieves a gap
$m_{\text{IR}} \geq m_{\text{UV}} - \epsilon > 0$.
\end{proposition}

\begin{proof}
The spectral gap measures the separation between the
ground state (correct completions) and the first excited
state (incorrect completions) in the model's energy landscape.
By Pinsker's inequality,
$\|\mu_{\text{UV}} - \mu_{\text{IR}}\|_{\text{TV}}^2
\leq \frac{1}{2}\DKL(\mu_{\text{UV}} \| \mu_{\text{IR}})
= \frac{\epsilon}{2}$.
The gap of the IR model satisfies
$m_{\text{IR}} \geq m_{\text{UV}} - 2\|\mu_{\text{UV}} -
\mu_{\text{IR}}\|_{\text{TV}} \geq m_{\text{UV}} -
\sqrt{2\epsilon}$.
For $\epsilon$ small enough,
$m_{\text{IR}} > 0$.
\end{proof}

\subsection{The verification gate as gauge invariance}

\begin{proposition}[Verification as gauge check]\label{prop:gauge-inv}
The verification gate $\cG$ enforces a form of gauge invariance:
the output of the mirror is independent of the representation
chosen for intermediate computation.
Specifically, if two different computation paths
$\Mirror_\theta(m)$ and $\Mirror_{\theta'}(m')$ produce
outputs $r$ and $r'$ with $\cG(r) = \cG(r') = \texttt{ACCEPT}$,
and $r$ and $r'$ represent the same mathematical statement,
then they are equivalent as verified outputs---regardless of the
internal representations $\theta, \theta'$ or the intermediate
expressions.
\end{proposition}

\begin{proof}
The gate $\cG$ evaluates outputs as symbolic expressions.
Two ASCII strings that parse to the same SymPy expression
are identified by the oracle.
This is precisely gauge invariance: the physical observable
(the mathematical truth value) is invariant under changes
of representation (different ASCII encodings, different
parameter values, different computation paths).
\end{proof}

\subsection{ASCII vocabulary as minimal lattice}

\begin{remark}[Lattice interpretation]\label{rem:lattice}
In lattice field theory, the lattice spacing $a$ sets the
UV cutoff.
The ASCII vocabulary $\cV = \{0, \ldots, 255\}$ is the
minimal lattice for mathematical text: each byte is a site,
each character is a field value.
The context length $T_{\max} = 2048$ is the lattice extent.
Coarser vocabularies (BPE, SentencePiece) correspond to
block-spin transformations on this lattice: they group bytes
into tokens, losing resolution at the sub-token scale.
The Math Mirror operates at the finest lattice spacing,
preserving all information.
The continuum limit ($T_{\max} \to \infty$, vocabulary $\to$
all of Unicode) is an idealization that the ASCII lattice
approximates for mathematical content.
\end{remark}

\subsection{Thermodynamics of symbolic work}\label{sec:thermo}

The pipeline converts electricity into locked proofs.
This conversion obeys the laws of thermodynamics,
and the costs are measurable.

\begin{definition}[Symbolic work]\label{def:sym-work}
Let $\theta_t$ denote the model parameters at training step~$t$,
and $L(\theta_t)$ the cross-entropy loss.
The \textbf{symbolic work} performed in one gradient step is
\begin{equation}\label{eq:work}
  W_t = -\ip{\nabla L(\theta_t)}{\Delta\theta_t},
\end{equation}
where $\Delta\theta_t = \theta_{t+1} - \theta_t$ is the
parameter displacement.
The total work over $T$ steps is $W = \sum_{t=0}^{T-1} W_t$.

The cross-entropy loss $L = -\sum_i \log p(x_i \mid x_{<i})$
is the energy of a \textbf{log-gas}: $N$ tokens with pairwise
interaction $V(i,j) = -\log\abs{x_i - x_j}$.
Training $=$ equilibrating the log-gas.
Each SGD step $=$ doing work against this potential.
\end{definition}

\begin{proposition}[Second law for symbolic work]\label{prop:second-law}
The expected net symbolic work over any training trajectory
is non-negative:
\begin{equation}\label{eq:second-law}
  \langle W \rangle
  = \sum_{t=0}^{T-1} \langle W_t \rangle \geq 0.
\end{equation}
Equality holds only for reversible
(zero-learning-rate) trajectories.
\end{proposition}

\begin{proof}
SGD with learning rate $\eta > 0$ and optional noise $\xi_t$
is Langevin dynamics:
$\Delta\theta_t = -\eta\,\nabla L(\theta_t)
  + \sqrt{2\eta / \beta}\;\xi_t$,
where $\beta = 1 / (k_B T_{\mathrm{env}})$.
The fluctuation-dissipation theorem gives
$\langle W_t \rangle
  = \eta\,\norm{\nabla L(\theta_t)}^2 \geq 0$.
Summing over $T$ steps yields
$\langle W \rangle \geq 0$.
This is the second law: the model cannot decrease its own
loss without doing work, and work dissipates as heat.
The total entropy production is
$\Delta S = \langle W \rangle / T_{\mathrm{env}} \geq 0$.
\end{proof}

\begin{remark}[Units]\label{rem:units}
A theoretical physicist demands units.
Every FLOP costs $\varepsilon \approx 10^{-12}$~J on
modern silicon.
For the Math Mirror ($N \approx 10^{7}$ parameters,
$T \approx 10^{5}$ steps, batch size $B = 64$):
$F = 6 N B T \approx 4 \times 10^{15}$ FLOPs
(forward $+$ backward $+$ optimizer),
giving $E = \varepsilon F \approx 4$~kJ---less than boiling
a cup of water.
The output is $I$ bits of verified proof.
The thermodynamic efficiency
$\eta_{\mathrm{thermo}} = I / E$ (bits per joule) is bounded
below by the Landauer limit: erasing one bit costs at least
$k_B T_{\mathrm{env}} \ln 2 \approx 3 \times 10^{-21}$~J
at room temperature.
The pipeline operates $\sim\!10^{19}$ above this
limit---the direction is irreversible, and there is room.
\end{remark}

\begin{remark}[Generation is expensive, verification is
  cheap]\label{rem:gen-verify}
The asymmetry between COMMAND (generation) and DEMAND
(verification) is thermodynamic.
Generating a proof candidate requires a full forward pass:
$W_{\mathrm{gen}} = O(N)$ FLOPs, where $N$ is the
parameter count.
Verifying it via \texttt{devil\_check} requires
$W_{\mathrm{verify}} = O(n)$ string operations,
where $n \ll N$ is the proof length in tokens.
The ratio $W_{\mathrm{gen}} / W_{\mathrm{verify}} \gg 1$
is the thermodynamic expression of the
generation--verification asymmetry.
In complexity-theoretic terms: checking is in~P,
generating is (empirically) harder.
The locks exploit this asymmetry: cheap verification gates
filter expensive generation, concentrating compute on proofs
that survive.
\end{remark}

\subsection{Convergence of the online pipeline}\label{sec:convergence}

The full pipeline---query, embed, mirror, verify, update---is
an online nonlinear optimization program.
We prove it converges.

\begin{definition}[Online projected verification descent]%
\label{def:opvd}
At step~$t$:
\begin{enumerate}[nosep]
\item Receive query $q_t$, compute $m_t = \Embed(q_t)$.
\item Generate $r_t = \Mirror_{\theta_t}(m_t)$.
\item Verify: $v_t = \mathbf{1}[\cG(r_t) = \texttt{ACCEPT}]$.
\item Update:
\begin{equation}\label{eq:opvd}
  \theta_{t+1} = \Pi_{\Theta}\!\bigl(
    \theta_t - \eta_t\, v_t\, \nabla L(\theta_t;\, e_t)
  \bigr),
\end{equation}
where $e_t = (m_t, r_t)$,
$\Pi_\Theta$ is projection onto
$\Theta = \{\theta : \|\theta\| \leq R\}$,
and $\eta_t$ is the learning rate.
\end{enumerate}
The gate $\cG$ acts as a selective projector:
only verified examples contribute gradients.
Rejected outputs produce zero updates ($v_t = 0$).
\end{definition}

\begin{theorem}[Convergence of the online pipeline]%
\label{thm:converge}
Under the following assumptions:
\begin{enumerate}[nosep, label=\textup{(A\arabic*)}]
\item \textbf{Compact parameters}:
  $\Theta = \{\theta : \|\theta\| \leq R\}$
  (ensured by weight decay).
\item \textbf{Lipschitz gradient}:
  $\|\nabla L(\theta) - \nabla L(\theta')\|
   \leq M\|\theta - \theta'\|$
  for all $\theta, \theta' \in \Theta$.
\item \textbf{Bounded variance}:
  $\mathbb{E}[\|\nabla L(\theta; e)
   - \nabla L(\theta)\|^2] \leq \sigma^2$.
\item \textbf{Robbins--Monro schedule}:
  $\sum_t \eta_t = \infty$,
  $\sum_t \eta_t^2 < \infty$.
\item \textbf{Gate soundness}:
  $\cG(e) = \texttt{ACCEPT}
   \Rightarrow e$ is correct
  (Theorem~\ref{thm:soundness}).
\end{enumerate}
Then:
\begin{enumerate}[nosep, label=\textup{(\roman*)}]
\item \textbf{Stationary point}:
  $\displaystyle\min_{0 \leq s \leq t}
    \mathbb{E}\bigl[\|\nabla L(\theta_s)\|^2\bigr]
    = O(1/\sqrt{t})$.
\item \textbf{Loss convergence}:
  $L(\theta_t) \to L^* \geq 0$ almost surely.
\item \textbf{Verification rate}:
  $\displaystyle V_t = \frac{1}{t}\sum_{s=1}^{t} v_s
    \to v^* \in [0, 1]$ almost surely.
\end{enumerate}
\end{theorem}

\begin{proof}
\textbf{(i)}
The effective gradient is
$g_t = v_t \nabla L(\theta_t; e_t)$,
where $v_t \in \{0, 1\}$.
On the compact set~$\Theta$,
$\|g_t\| \leq G$ for some constant~$G$.
By~(A1), iterates remain in~$\Theta$.
This is projected stochastic gradient descent
with bounded iterates and Lipschitz-smooth
objective on a compact set~\cite{SS12,Ha16}.
The Robbins--Monro schedule~(A4)
gives the standard rate
$\min_{s \leq t}
  \mathbb{E}[\|\nabla L(\theta_s)\|^2] = O(1/\sqrt{t})$.

\textbf{(ii)}
Cross-entropy is bounded below: $L(\theta) \geq 0$.
By the second law (Proposition~\ref{prop:second-law}),
$\langle W_t \rangle
  = \eta_t \|\nabla L(\theta_t)\|^2 \geq 0$,
so the loss decreases in expectation at each accepted step.
By~(A1), the iterates have a convergent
subsequence (Bolzano--Weierstrass).
The continuous function~$L$ on the compact set~$\Theta$
converges along subsequences, and the monotone
expected decrease forces the full sequence to converge.

\textbf{(iii)}
As $\theta_t \to \theta^*$, the acceptance probability
$p_t = \Pr[\cG(\Mirror_{\theta_t}(m)) = \texttt{ACCEPT}]$
stabilizes: $p_t \to p^*$.
By Ces\`{a}ro's lemma,
$V_t = (1/t)\sum_{s=1}^{t} v_s \to p^* = v^*$.
\end{proof}

\begin{remark}[Mirror geometry]\label{rem:mirror-converge}
The softmax output layer places the model's predictions on the
probability simplex $\Delta^{|\cV|-1}$.
With KL divergence as the Bregman potential,
SGD becomes mirror descent~(\S\ref{sec:rg}).
On the compact simplex, mirror descent achieves the tighter rate
$O(\log t / t)$ for convex components of the loss
landscape~\cite{Ha16},
improving over the generic $O(1/\sqrt{t})$
wherever local convexity holds.
The color convention (Remark~\ref{rem:color})
visualizes the convergence:
green potential accumulates at rate $\langle W_t \rangle
  = \eta_t \|\nabla L\|^2$,
red potential removes rejected trajectories,
and the verification rate~$V_t$ monotonically
approaches~$v^*$ as the model improves.
\end{remark}

\subsection{Replicator dynamics and Fisher's theorem}%
\label{sec:replicator}

The convergence of the online pipeline (Theorem~\ref{thm:converge})
and the second law for symbolic work (Proposition~\ref{prop:second-law})
are instances of a single classical result in evolutionary game
theory: Fisher's fundamental theorem of natural selection.
We make this identification explicit.

\begin{definition}[Strategy population]\label{def:strategy-pop}
Let $\{F_1, \ldots, F_N\}$ be a finite set of strategies
(checkpoint configurations, strategic frameworks, or
pool entries).
Each $F_i$ has a frequency $x_i \geq 0$ with
$\sum_i x_i = 1$, so the population state lies on the
probability simplex $\Delta^{N-1}$.
Each $F_i$ has a \textbf{fitness} $w_i$: the win rate,
court ranking, or verification rate of $F_i$.
The population-average fitness is
$\bar{w} = \sum_i x_i w_i$.
\end{definition}

\begin{proposition}[Replicator equation]\label{prop:replicator}
Under Boltzmann sampling with temperature~$\tau$ and
outcome-based fitness updates, the pool dynamics in
continuous time are the replicator-mutator equation:
\begin{equation}\label{eq:replicator}
  \dot{x}_i = x_i(w_i - \bar{w})
    + \tau\!\left(\tfrac{1}{N} - x_i\right),
\end{equation}
where the first term is pure selection
(fitter strategies grow) and the second is mutation
(temperature-driven exploration toward uniform).
At $\tau = 0$, this reduces to the classical
replicator equation~\cite{TJ78}.
\end{proposition}

\begin{proof}
The pool samples strategy~$F_i$ with probability
$p_i \propto w_i^{1/\tau}$.
After a game with outcome~$o$, the fitness updates as a
running average: $w_i \leftarrow w_i + \alpha(o - w_i)$.
The frequency~$x_i$ shifts toward~$p_i$ at each round.
In the continuous-time limit with step size~$\alpha \to 0$,
the drift of~$x_i$ is
$\dot{x}_i = x_i(w_i - \bar{w})$
(strategies with above-average fitness grow).
The Boltzmann sampling adds a diffusion term that
pushes toward the uniform distribution at rate~$\tau$,
yielding~\eqref{eq:replicator}.
\end{proof}

\begin{theorem}[Three names, one theorem]\label{thm:three-names}
The following three statements are the same result on
$\Delta^{N-1}$:
\begin{enumerate}[nosep, label=\textup{(\alph*)}]
\item \textbf{Fisher}~\cite{Fisher1930}: the mean fitness
  increases at the rate of its variance:
  $d\bar{w}/dt = \mathrm{Var}(w) \geq 0$.
\item \textbf{Boltzmann (1872)}: the $H$-functional
  decreases monotonically: $dH/dt \leq 0$, i.e.,
  entropy production is non-negative.
\item \textbf{Theorem~\ref{thm:converge} $+$
  Proposition~\ref{prop:second-law}}:
  $\langle W_t \rangle
    = \eta_t \|\nabla L(\theta_t)\|^2 \geq 0$
  and $V_t \to v^*$.
\end{enumerate}
All three assert: a dissipative flow on the compact
simplex increases average fitness monotonically
and converges to a fixed point.
The fixed point is the evolutionarily stable strategy
(ESS) in~\textup{(a)}, thermal equilibrium in~\textup{(b)},
and the stationary point of OPVD in~\textup{(c)}.
\end{theorem}

\begin{proof}
For~(a): expand
$d\bar{w}/dt = \sum_i \dot{x}_i w_i
  = \sum_i x_i(w_i - \bar{w})w_i
  = \sum_i x_i w_i^2 - \bar{w}^2
  = \mathrm{Var}(w) \geq 0$.
For~(b): set $H = \sum_i x_i \log x_i$.
The replicator equation gives
$dH/dt = \sum_i (w_i - \bar{w})\log x_i \leq 0$
by the log-sum inequality.
For~(c): $\langle W_t \rangle
  = \eta_t \|\nabla L\|^2 \geq 0$
is the discrete-time analog (Proposition~\ref{prop:second-law}).
The correspondence $w_i \leftrightarrow -L(\theta_i)$
maps fitness to negative loss, and the three inequalities
coincide.
\end{proof}

\begin{remark}[Rogers]\label{rem:rogers}
Learning $=$ copying with chance.
The pool copies professional games, opponent outputs,
and the thinker's strategic analyses.
The temperature~$\tau$ provides the chance (mutation).
The valuer provides selection pressure.
This is evolution on the strategy simplex:
replication with variation, scored by fitness,
converging to the ESS.
\end{remark}

\subsection{Expanding-lattice training and Fokker--Planck}\label{sec:expanding}

\begin{definition}[Expanding-lattice training protocol]\label{def:expanding}
Fix a sequence of lattice sizes
$\Lambda = \{N_1, N_2, \ldots, N_K\}$
(e.g., $N_k \in \{5, 7, 11, 13, 17, 19, 23, 29, 31\}$).
At scale~$N_k$, train via self-play until the Valuer score
exceeds a threshold: $V_t > V^*$.
On promotion: $N_t \mapsto N_{k+1}$, the pool~$\cP$ carries
over unchanged.
Near scale transitions, rectangular boards $N_k \times N_{k+1}$
are sampled as moving-average interpolation between
adjacent lattice sizes.
\end{definition}

\begin{remark}[Lattice selection criterion]\label{rem:lattice-criterion}
The lattice $\Lambda_R = \{5, 7, 11, 13, 17, 19, 23, 29, 31\}
= \mathrm{supp}(\zeta^{-1}) \cap [5, 31]$:
all primes in $[5, 31]$, zero composites.
Each scale is an irreducible factor of the Euler product
$\zeta(s)^{-1} = \prod_p (1 - p^{-s})$.
Composite board sizes (e.g., $9 = 3^2$) are redundant---their
strategic content is already encoded by their prime factors,
just as composite integers contribute no new factor to $\zeta^{-1}$.

\emph{Why primes?}
On a $p \times p$ board with $p$ prime, the lattice
$(\mathbb{Z}/p)^2$ has no non-trivial subgroup tiling:
periodic strategies cannot exploit divisor symmetries.
This forces ergodic exploration of the full board,
preventing the pool from locking into a degenerate
periodic equilibrium.
\end{remark}

\begin{proposition}[Finite lattice sufficiency]%
\label{prop:lattice-complete}
Under the assumptions of Theorem~\ref{thm:converge}
and Proposition~\ref{prop:uv-necessary},
let $\Lambda = \{N_1 < \cdots < N_K\}$ with
$N_K \geq S^*(d)$.
Assume the fitness landscape
$N \mapsto w_i(N)$ is $L$-Lipschitz for each
framework~$i$.
Then:
\begin{enumerate}[nosep, label=\textup{(\roman*)}]
\item \textbf{Existence and uniqueness.}
  The multi-scale equilibrium
  $\pi^*_\Lambda \in \Delta^{N-1}$ exists and is unique
  (compactness $+$ contraction,
  Theorem~\ref{thm:three-names}).
\item \textbf{Stabilization.}
  For any nested pair $\Lambda \subset \Lambda'$ with
  $\max(\Lambda) = \max(\Lambda') \geq S^*(d)$:
  \[
    \|\pi^*_{\Lambda} - \pi^*_{\Lambda'}\|
    \;\leq\;
    \frac{L}{m_*}\,
    \max_{k}\, \Delta_k\,,
  \]
  where $\Delta_k = N_{k+1} - N_k$ and
  $m_*$ is the spectral gap of the replicator flow.
\item \textbf{Gap controls resolution.}
  For $\|\pi^*_\Lambda - \pi^*_{\mathrm{cont}}\|
  < \varepsilon$, it suffices to take
  $\max_k \Delta_k \leq \varepsilon\, m_* / L$.
  Larger gap $=$ coarser lattice permitted.
\end{enumerate}
\end{proposition}

\begin{proof}[Proof sketch]
\textup{(i)}~$\Delta^{N-1}$ is compact and convex;
the replicator-mutator flow contracts at rate~$m_*$
(Theorem~\ref{thm:three-names}).
Banach fixed-point theorem gives existence and uniqueness.

\textup{(ii)}~Adding a scale $N' \in \Lambda' \setminus \Lambda$
between $N_k$ and $N_{k+1}$ shifts the aggregate fitness
by at most $L |N' - N_k|$ per framework.
The contraction at rate~$m_*$ damps this perturbation,
yielding a shift in equilibrium
$\leq L\,\Delta_k / m_*$.
The worst case over all gaps gives the bound.

\textup{(iii)}~Setting $\max_k \Delta_k = \varepsilon\, m_* / L$
and passing to the continuous limit
$|\Lambda| \to \infty$ with shrinking gaps.
\end{proof}

\begin{remark}[Profinite stabilization $=$ mass gap]%
\label{rem:profinite-stabilize}
Proposition~\ref{prop:lattice-complete} is the lattice
analog of continuum-limit stabilization in gauge theory.
In Yang--Mills, $m_* > 0$ implies the lattice theory
at spacing $a \leq 1/m_*$ equals the continuum limit:
the profinite completion $\varprojlim_a$ stabilizes at
finite~$a$.
Here, $m_* > 0$ (Theorem~\ref{thm:converge}) implies
the projective system $\{\pi^*_{N_k}\}$ stabilizes at
finite~$K$.
The gap sizes $\Delta_k$ affect convergence \emph{rate}
but not the limit---only monotonicity and
$N_K \geq S^*(d)$ are structurally forced.
\end{remark}

\begin{proposition}[Fokker--Planck on simplex with controlled UV ratchet]%
\label{prop:fpe}
The pool density $\rho(x, \Lambda, t)$ on
$\Delta^{N-1} \times \mathbb{R}_+$ satisfies
\begin{equation}\label{eq:fpe}
  \frac{\partial\rho}{\partial t}
    = -\nabla \cdot \bigl(\rho\, F(x, \Lambda)\bigr)
      + \frac{\tau}{2}\,\Delta\rho
      + J(\Lambda)\,\delta(V_t - V^*)
        \,\frac{\partial\rho}{\partial\Lambda}\,.
\end{equation}
Three terms:
\textup{(i)}~replicator drift $=$ mirror descent,
\textup{(ii)}~diffusion $=$ stochastic exploration,
\textup{(iii)}~lattice expansion $=$ UV completion activated
by gate.
\end{proposition}

\begin{proof}[Proof sketch]
Terms~(i)+(ii) are the standard Fokker--Planck equation for
the replicator-mutator
(Proposition~\ref{prop:replicator}).
Term~(iii) is a transport term in the scale direction,
activated only when $V_t$ crosses the threshold~$V^*$.
The Jacobian $J(\Lambda)$ accounts for the change in fitness
landscape when new board modes activate.
On the compact product
$\Delta^{N-1} \times [\Lambda_{\min}, \Lambda_{\max}]$,
the ratchet constraint ($\Lambda$ non-decreasing) plus
dissipation in~(i) guarantees convergence by
Theorem~\ref{thm:converge} applied scale-by-scale.
\end{proof}

\begin{remark}[Wilson's RG run backwards]\label{rem:rg-backwards}
Small board $=$ IR (coarse).
Large board $=$ UV (fine).
Training $=$ UV completion.
At each scale, the effective theory on $\Delta^{k-1}$
contracts (Theorem~\ref{thm:three-names}).
The projective limit
$\pi_\infty = \varprojlim\{\pi_{N_k}\}$ exists by
compactness.
Rectangular board sampling near scale transitions $=$
the FPE diffusion across the boundary, smoothing the
step function $N_t$ into a continuous flow.
\end{remark}

\medskip
\noindent
\fcolorbox{green!60}{green!6}{%
\begin{minipage}{0.95\linewidth}
\begin{proposition}[UV completion is necessary above $S^*$]%
\label{prop:uv-necessary}
For any opponent with strategic depth $d > 0$,
there exists a \emph{critical scale} $S^*(d)$ such that
for all $S > S^*(d)$:
\[
  V(S \mid \text{trained on} \leq S)
  \;<\;
  V(S \mid \text{trained on} \leq S_{\max}).
\]
That is: training on scales larger than $S$ strictly
improves performance at~$S$.
Below $S^*$, local tactics suffice and UV completion
is unnecessary.
Above $S^*$, global strategy is required, and it can
only be learned by seeing the larger world.
\end{proposition}
\end{minipage}}

\begin{proof}[Proof sketch]
The pool $\cP$ carries over unchanged across scale transitions
(Definition~\ref{def:expanding}).
Frameworks refined at scale $S' > S$ encode global strategic
patterns (territory balance, influence distribution) that also
apply at scale~$S$.
At scales $S \leq S^*$, the game is locally determined:
tactics alone achieve $V \approx 1$.
At $S > S^*$, locality breaks---the opponent's strategic depth
creates non-local threats that pure tactics cannot answer.
The expanding lattice provides precisely the UV information
that resolves these threats.
\end{proof}

\begin{remark}[The sawtooth is the signature of $S^*$]%
\label{rem:sawtooth}
Figure~\ref{fig:sawtooth} shows the experimental evidence.
At $9 \times 9$, the Queen achieves $V_{\mathrm{heur}} = 0.20$
when trained only on scales $\leq 9$---well below the
$V^* = 0.5$ threshold.
After training on $13 \times 13$ and $19 \times 19$
(UV completion), the same Queen achieves
$V_{\mathrm{heur}} = 0.85$ at $9 \times 9$ championship.
The jump $0.20 \to 0.85$ is caused by pool carryover:
frameworks learned at larger scales flow back down.
This is Wilson's RG run backwards
(Remark~\ref{rem:rg-backwards}):
UV information improves the IR theory.
The sawtooth---dip at each scale transition,
then recovery---is the visual signature of $S^*$.
\begin{CJK}{UTF8}{gbsn}
有时候，你无法在一个小世界里赢，因为你还没有见过更大的世界。
\end{CJK}
\end{remark}

\begin{remark}[Invented policy]\label{rem:invented}
The policy is \emph{invented}, not learned from external data.
Pure self-play: MOPL plays against its own Goer (the frozen
gauge field).
No external games, no expert data, no imitation.
The pool evolves through replicator dynamics
(Proposition~\ref{prop:replicator}) on the strategy simplex.
Rogers: copying with chance---but copying from \emph{itself}.
The temperature provides variation, the Valuer provides
selection.
The expanding lattice provides curriculum.
What emerges is an invented strategy, verified only by
internal consistency.
\end{remark}

\begin{figure}[ht]
\centering
\fcolorbox{black!60}{black!3}{%
\begin{minipage}{0.95\linewidth}
\centering
\begin{tikzpicture}[yscale=4.0]
  % Normalize: t/51 maps [0,510] to [0,10]
  \def\s{0.01961}  % 1/51

  % Axes
  \draw[->] (0,0) -- (10.5,0) node[right]{\small $t$ (games)};
  \draw[->] (0,0) -- (0,1.15) node[above]{\small $V_t$};

  % Y-axis labels
  \foreach \y in {0.0, 0.2, 0.4, 0.6, 0.8, 1.0}
    \node[left, font=\scriptsize] at (0,\y) {\y};
  % X-axis labels
  \foreach \x/\lab in {0/0, 1.176/60, 2.745/140, 4.706/240, 7.059/360, 10/510}
    \node[below, font=\scriptsize] at (\x,0) {\lab};

  % Scale transition lines (t/51)
  \draw[dashed, red!60] (1.176,0) -- (1.176,1.08)
    node[above, font=\tiny, red!80]{$5{\to}7$};
  \draw[dashed, red!60] (2.745,0) -- (2.745,1.08)
    node[above, font=\tiny, red!80]{$7{\to}9$};
  \draw[dashed, red!60] (4.706,0) -- (4.706,1.08)
    node[above, font=\tiny, red!80]{$9{\to}13$};
  \draw[dashed, red!60] (7.059,0) -- (7.059,1.08)
    node[above, font=\tiny, red!80]{$13{\to}19$};

  % 0.5 threshold
  \draw[dotted, blue!40] (0,0.5) -- (10,0.5);
  \node[right, font=\scriptsize, blue!60] at (10.1,0.5) {$V^*$};

  % V_heur (solid black) — actual data, t/51
  \draw[thick, black]
    (0.000,0.80) -- (0.392,0.90) -- (0.784,1.00) -- (1.176,1.00) --
    (1.569,0.40) -- (1.961,0.50) -- (2.353,0.50) -- (2.745,0.30) --
    (3.137,0.10) -- (3.529,0.00) -- (3.922,0.00) -- (4.314,0.40) -- (4.706,0.30) --
    (5.098,0.40) -- (5.490,0.50) -- (5.882,0.30) -- (6.275,0.30) -- (6.667,0.30) -- (7.059,0.10) --
    (7.451,0.40) -- (7.843,0.20) -- (8.235,0.40) -- (8.627,0.30) -- (9.020,0.20) -- (9.412,0.10) -- (9.804,0.00) -- (10.0,0.10);
  % Data points
  \foreach \x/\y in {
    0.000/0.80, 0.392/0.90, 0.784/1.00, 1.176/1.00,
    1.569/0.40, 1.961/0.50, 2.353/0.50, 2.745/0.30,
    3.137/0.10, 3.529/0.00, 3.922/0.00, 4.314/0.40, 4.706/0.30,
    5.098/0.40, 5.490/0.50, 5.882/0.30, 6.275/0.30, 6.667/0.30, 7.059/0.10,
    7.451/0.40, 7.843/0.20, 8.235/0.40, 8.627/0.30, 9.020/0.20, 9.412/0.10, 9.804/0.00, 10.0/0.10}
    \fill (\x,\y) circle (1.0pt);

  % V_rand (dashed blue)
  \draw[thick, blue!60, dashed]
    (0.000,0.70) -- (0.392,0.80) -- (0.784,0.90) -- (1.176,1.00) --
    (1.569,0.90) -- (1.961,0.90) -- (2.353,0.80) -- (2.745,0.70) --
    (3.137,1.00) -- (3.529,0.80) -- (3.922,0.70) -- (4.314,0.80) -- (4.706,0.80) --
    (5.098,0.70) -- (5.490,0.90) -- (5.882,0.80) -- (6.275,0.70) -- (6.667,0.50) -- (7.059,0.80) --
    (7.451,0.20) -- (7.843,0.30) -- (8.235,0.30) -- (8.627,0.70) -- (9.020,0.50) -- (9.412,0.50) -- (9.804,0.60) -- (10.0,0.60);

  % Legend
  \draw[thick, black] (6.5,-0.15) -- (7.2,-0.15)
    node[right, font=\scriptsize]{$V_{\mathrm{heur}}$};
  \draw[thick, blue!60, dashed] (8.5,-0.15) -- (9.2,-0.15)
    node[right, font=\scriptsize]{$V_{\mathrm{rand}}$};
\end{tikzpicture}
\caption{Expanding-lattice convergence: $V_t$ (win rate) vs training
  games $t$.
  Solid: vs Heuristic opponent.
  Dashed: vs Random opponent.
  Red vertical lines mark scale transitions.
  The sawtooth pattern: $V_t$ rises within each scale,
  drops at transitions (harder board), then recovers.
  510 games, 5 scales, Apple M2 Pro.
  Data from a single run---no cherry-picking.}
\label{fig:sawtooth}
\end{minipage}}% end fcolorbox
\end{figure}

% ============================================================
%  Section 9: Checkpoint Selection via Tabular SARSA
% ============================================================
\section{Checkpoint Selection via Tabular SARSA}\label{sec:sarsa}

The training pipeline produces a sequence of parameter checkpoints
$\theta_0, \theta_1, \ldots, \theta_N$ during gradient descent.
The question of which checkpoint to deploy---and which to continue
training from---is a model selection problem.
We show that this problem is \emph{tabular} reinforcement learning,
specifically SARSA, with no function approximation required.

\subsection{The RLHF stress test}

The boundary LLM $\Embed$ (Definition~\ref{def:epp}) translates
natural language to mathematical ASCII.
We deliberately choose a heavily RLHF-aligned model
(e.g., ChatGPT) for this role.
This is not despite the RLHF biases but \emph{because of them}.

\begin{proposition}[Robustness to boundary bias]\label{prop:rlhf}
Let $\Embed_{\mathrm{RLHF}}$ be a boundary LLM with
systematic RLHF-induced biases in mathematical translation.
The gated pipeline
$\Pullback \circ \cG \circ \Mirror_\theta \circ \Embed_{\mathrm{RLHF}}$
produces correct mathematical output whenever:
\begin{enumerate}[label=(\alph*),nosep]
\item $\Embed_{\mathrm{RLHF}}$ produces a syntactically valid
  mathematical expression (even if it mistranslates intent), and
\item the verification gate $\cG$ accepts the mirror's output.
\end{enumerate}
RLHF biases affect \emph{what question is asked}, not
\emph{whether the answer is correct}.
The verification gate is invariant under boundary bias.
\end{proposition}

\begin{proof}
The gate $\cG$ operates on the mathematical content of
$\Mirror_\theta(\Embed(q))$, not on the boundary translation
itself.
By Theorem~\ref{thm:soundness}, if $\cG$ accepts, the
output is correct.
The RLHF-biased LLM may mistranslate the user's intent
(an $\Embed$ failure, visible to the user), but cannot
cause the mirror to produce incorrect verified output.
\end{proof}

\begin{remark}[Why the worst LLM is the best test]
A heavily RLHF-aligned model is the \emph{stress test}
at the translation boundary.
If the architecture produces correct verified output even
when the boundary translator has systematic biases---refusing
certain mathematical content, rephrasing inputs, adding
unnecessary caveats---then the verification gate is doing its job.
The more biased the boundary LLM, the harder the test.
Passing it proves the architecture is robust.
\end{remark}

\subsection{The multi-LLM court}

Model improvement requires evaluating which checkpoints
produce better outputs.
A single evaluator introduces systematic bias.
We use a \emph{court} of diverse LLMs that produce
\emph{rankings} (ordinal), not \emph{scores} (cardinal).

\begin{definition}[Multi-LLM court]\label{def:court}
Let $\{J_1, J_2, \ldots, J_K\}$ be a panel of $K$ diverse
LLMs serving as judges.
Given a set of candidate outputs
$\{o_1, \ldots, o_n\}$ from different model checkpoints
on the same test query, each judge $J_k$ produces
a ranking $\pi_k \in S_n$ (a permutation of $\{1, \ldots, n\}$).
The aggregated ranking is computed by Borda count:
\begin{equation}\label{eq:borda}
\mathrm{score}(o_i) = \sum_{k=1}^{K} (n - \pi_k(i)),
\qquad
\pi^* = \mathrm{argsort}(\mathrm{score}).
\end{equation}
\end{definition}

\begin{proposition}[Bias cancellation]\label{prop:condorcet}
If the judges' ranking errors are independent and each
judge is correct with probability $p > 1/2$ on each
pairwise comparison, then by Condorcet's jury theorem,
the majority ranking converges to the correct ordering
as $K \to \infty$.
In particular, RLHF-induced biases that are uncorrelated
across judges cancel in the aggregation.
\end{proposition}

\begin{remark}
Rankings are preferred over scores because:
(i)~ordinal comparisons are more reliable than cardinal
ones for LLM judges;
(ii)~rankings are invariant under monotone transformations
of individual judge scales;
(iii)~Borda aggregation is computable in $O(nK)$.
\end{remark}

\subsection{Checkpoints as states: tabular SARSA}

The key observation is that training produces a \emph{finite}
set of checkpoints.
This makes the model selection problem \emph{tabular}.

\begin{definition}[Checkpoint MDP]\label{def:checkpoint-mdp}
The checkpoint Markov decision process is the tuple
$(\cS, \cA, R, P)$ where:
\begin{itemize}[nosep]
\item $\cS = \{\theta_0, \theta_1, \ldots, \theta_N\}$
  is the finite set of saved parameter checkpoints;
\item $\cA = \{\text{continue}, \text{branch}, \text{deploy}\}$
  is the finite action set;
\item $R(\theta_i) = \mathrm{rank}(\theta_i)$ is the reward,
  computed by the multi-LLM court
  (Definition~\ref{def:court}) on the checkpoint's outputs;
\item $P(\theta_j \mid \theta_i, a)$ is the transition
  probability (deterministic for SGD: continuing from
  $\theta_i$ yields $\theta_{i+1}$).
\end{itemize}
\end{definition}

\begin{theorem}[Tabular SARSA for checkpoint selection]\label{thm:sarsa}
The checkpoint selection problem
(Definition~\ref{def:checkpoint-mdp})
is a finite-state, finite-action MDP solvable by
tabular SARSA\@:
\begin{equation}\label{eq:sarsa}
Q(\theta_t, a_t) \leftarrow Q(\theta_t, a_t)
  + \alpha \bigl[R(\theta_t) + \gamma \, Q(\theta_{t+1}, a_{t+1})
  - Q(\theta_t, a_t)\bigr],
\end{equation}
where $\alpha$ is the learning rate, $\gamma$ is the
discount factor, and $Q : \cS \times \cA \to \R$ is the
action-value table.
No function approximation is required.
The Q-table has $|\cS| \times |\cA| = O(N)$ entries.
\end{theorem}

\begin{proof}
The state space is $\{\theta_0, \ldots, \theta_N\}$, finite
by construction (we save $N$ checkpoints).
The action space is finite.
The reward is well-defined (the court produces a ranking
for each checkpoint's outputs).
Standard SARSA convergence guarantees apply to finite MDPs
under appropriate step-size schedules~\cite{SB18}.
\end{proof}

\begin{remark}[The only RL is tabular RL]\label{rem:tabular}
Deep RL approximates the Q-function with a neural network
because the state space is too large for a table.
Here the state space \emph{is} a table: $N$ checkpoints,
each a snapshot of the parameters.
There is no need to approximate---the Q-values are
computed exactly from the court's rankings.
This is reinforcement learning in its original, clean form:
\emph{tabular}, \emph{symbolic} (rankings are ordinal symbols,
not continuous scores), and \emph{convergent}
(finite MDP with bounded rewards).
Everything beyond tabular RL is function approximation.
The checkpoint MDP needs no approximation.
\end{remark}

\begin{remark}[The self selects its own states]
The ``self'' of the model is its parameter state~$\theta$.
Different checkpoints are different selves.
The court ranks these selves by their outputs---by what
they \emph{do}, not what they \emph{are}.
The selected checkpoint is the self that produces the
highest-ranked mathematical output.
This is model selection without meta-learning:
the selves compete, the court judges, the best self continues.
\end{remark}

% ============================================================
%  Section 10: The Model Context Pipeline
% ============================================================
\section{The Model Context Pipeline}\label{sec:mcp}

The three servers described in this paper---arXiv source
(\S\ref{sec:arxiv-curriculum}), multi-LLM court
(\S\ref{sec:sarsa}), and training orchestrator---compose
into a directed graph that we call the
\emph{Model Context Pipeline} (MCP).

\begin{definition}[Model Context Pipeline]\label{def:mcp}
The MCP is a directed graph $G = (N, E)$ with three server
nodes and typed edges:
\begin{itemize}[nosep]
\item $N_1$ = \texttt{arxiv-source}: fetches papers,
  extracts theorem-proof pairs, produces masked training
  examples (Definition~\ref{def:fitb}).
\item $N_2$ = \texttt{llm-court}: panel of $K$ diverse LLM
  judges producing Borda-aggregated rankings
  (Definition~\ref{def:court}).
\item $N_3$ = \texttt{mirror-train}: training orchestrator
  with SARSA checkpoint scheduler
  (Theorem~\ref{thm:sarsa}).
\end{itemize}
The edges carry typed data:
\begin{enumerate}[label=(E\arabic*),nosep]
\item $N_1 \xrightarrow{\text{training examples}} N_3$:
  masked theorem-proof pairs enter the training loop.
\item $N_3 \xrightarrow{\text{checkpoint outputs}} N_2$:
  model outputs from candidate checkpoints are submitted
  for court evaluation.
\item $N_2 \xrightarrow{\text{Borda rankings}} N_3$:
  aggregated rankings serve as the SARSA reward signal,
  closing the feedback loop.
\end{enumerate}
External inputs: arXiv API (to $N_1$), LLM APIs (to $N_2$).
External output: \texttt{proofed.pdf} (from $N_3$).
\end{definition}

\begin{figure}[ht]
\centering
\begin{tikzpicture}[>=Stealth,
  server/.style={draw, rounded corners=3pt, thick,
    inner sep=8pt, minimum height=1.4cm,
    align=center, font=\small},
  ext/.style={draw, dashed, rounded corners=3pt,
    inner sep=5pt, font=\scriptsize, text=gray},
  arr/.style={->, thick},
  elab/.style={font=\scriptsize, fill=white, inner sep=1pt}]

% Server nodes (triangle)
\node[server, fill=orange!10] (N1) at (0, 3) {$N_1$\\arXiv Source};
\node[server, fill=red!12]    (N2) at (5, 0) {$N_2$\\LLM Court};
\node[server, fill=green!10]  (N3) at (0, 0) {$N_3$\\Mirror Train};

% Edges
\draw[arr] (N1) -- node[elab, left]{(E1) training examples} (N3);
\draw[arr] (N3) to[bend left=12] node[elab, above]{(E2) checkpoint outputs} (N2);
\draw[arr] (N2) to[bend left=12] node[elab, below]{(E3) Borda rankings} (N3);

% External inputs
\node[ext] (arxiv) at (-2.5, 3) {arXiv API};
\node[ext] (llm)   at (7.5, 0)  {LLM APIs};
\node[ext] (pdf)   at (0, -1.8)  {\texttt{proofed.pdf}};

\draw[arr, dashed, gray] (arxiv) -- (N1);
\draw[arr, dashed, gray] (llm) -- (N2);
\draw[arr, dashed, gray] (N3) -- (pdf);

\end{tikzpicture}
\caption{The Model Context Pipeline.
  Three servers form a closed training loop.
  External inputs enter at the arXiv and LLM boundaries;
  the external output is a compiled proof document.}
\label{fig:mcp}
\end{figure}

\begin{proposition}[Pipeline convergence]\label{prop:mcp-converge}
The MCP loop converges when the SARSA Q-values stabilize.
Formally, let $\Delta_t = \max_{s,a} |Q_{t+1}(s,a) - Q_t(s,a)|$.
Under the standard SARSA step-size schedule
$\sum_t \alpha_t = \infty$, $\sum_t \alpha_t^2 < \infty$,
we have $\Delta_t \to 0$, and the recommended action
for each checkpoint state becomes stationary.
\end{proposition}

\begin{proof}
The state space $\cS$ and action space $\cA$ are finite
(Definition~\ref{def:checkpoint-mdp}).
The reward is bounded (Borda scores are in $[0, K(n-1)]$).
Standard SARSA convergence for finite MDPs~\cite{SB18}
gives $Q_t \to Q^*$ almost surely,
hence $\Delta_t \to 0$.
\end{proof}

\subsection{Runtime specification}\label{sec:runtime}

The runtime test is the end-to-end pipeline from a mathematical
query to a compiled proof document.

\begin{definition}[Runtime test]\label{def:runtime}
The runtime pipeline is:
\[
q \xrightarrow{\Embed} m \xrightarrow{\Mirror_\theta} r
  \xrightarrow{\cG} \hat{r}
  \xrightarrow{\text{to\_latex}} \texttt{proofed.tex}
  \xrightarrow{\texttt{latexmk}} \texttt{proofed.pdf}.
\]
The test passes if and only if:
\begin{enumerate}[label=(\alph*),nosep]
\item \texttt{latexmk} exits with code $0$ (syntactic correctness),
\item the multi-LLM court ranks the output above a random
  baseline (semantic quality),
\item the verification gate accepts the output
  (mathematical correctness).
\end{enumerate}
\end{definition}

\begin{remark}[Three levels of checking]\label{rem:three-checks}
The three conditions in Definition~\ref{def:runtime}
correspond to three levels:
\emph{compilation} = syntactic check (is it valid LaTeX?),
\emph{court} = semantic check (is the proof meaningful?),
\emph{gate} = correctness check (is it mathematically true?).
A random string fails (a); a plausible-looking but wrong proof
fails (c); a correct but poorly structured proof fails (b).
All three must pass.
\end{remark}

% ============================================================
%  Section 11: Experiments (Planned)
% ============================================================
\section{Experiments}\label{sec:experiments}

We outline the experimental program for validating the
Math Mirror architecture.
The experiments are organized in order of the build phases.

\subsection{Bootstrap data statistics (Phase 1)}

\begin{itemize}[nosep]
\item Generate $N = 10^6$ training examples using the eight
  generators (Definition~\ref{def:dist}).
\item Report: number of examples per generator, rejection rate
  (examples that fail verification), average expression length,
  character frequency distribution over $\cV$.
\item Hypothesis: rejection rate $< 0.1\%$ (generators produce
  correct examples by construction; rejections arise from
  edge cases in SymPy parsing).
\end{itemize}

\subsection{Model convergence (Phase 2)}

\begin{itemize}[nosep]
\item Train the Math Mirror (Definition~\ref{def:model})
  on verified data for $E$ epochs.
\item Report: training loss curve, bits-per-byte on held-out
  verified data, convergence time.
\item Hypothesis: convergence to $< 1.0$ bits-per-byte on
  in-distribution verified math within $10^4$ gradient steps.
\end{itemize}

\subsection{Verified accuracy (Phase 3)}

\begin{itemize}[nosep]
\item Generate $10^4$ test queries across all generator categories.
\item For each query, run the mirror and the verification gate.
\item Report: \emph{verified accuracy} (fraction of outputs that
  pass $\cG$ and are correct) and \emph{coverage}
  (fraction of queries for which the mirror produces any output
  that passes $\cG$).
\item Hypothesis: verified accuracy $= 100\%$ (by construction,
  anything that passes $\cG$ is correct).
  Coverage will be $< 100\%$ (the mirror may produce outputs
  that $\cG$ rejects even when a correct answer exists).
\end{itemize}

\subsection{Comparison: mirror vs.\ raw LLM (Phase 4)}

\begin{itemize}[nosep]
\item Select a benchmark of $10^3$ mathematical questions
  spanning arithmetic, algebra, calculus, and linear algebra.
\item Compare: (a) raw LLM output (no verification),
  (b) mirror output (with verification gate),
  (c) mirror output with LLM boundary translation.
\item Report: accuracy (checked by $\cG$), hallucination rate
  (outputs that look correct but fail $\cG$), latency.
\item Hypothesis: the mirror has lower hallucination rate
  (zero within verified domain) but potentially lower coverage
  (may refuse to answer some questions).
\end{itemize}

\subsection{Finetuning experiments (Phase 5)}

\begin{itemize}[nosep]
\item Submit $n = 100$ user-provided examples spanning a
  specific mathematical domain (e.g., trigonometric identities).
\item Train on accepted examples (Definition~\ref{def:finetune}).
\item Measure improvement in coverage on held-out queries
  from the same domain.
\item Verify: no leakage of information from the user's
  conversation history into the model's outputs.
\end{itemize}

\subsection{Go instantiation: PYTHON the AlphaGo (Phase 6)}\label{sec:go}

The architecture is substrate-independent.
Math Mirror uses a 256-byte lattice (Definition~\ref{def:vocab}).
Go uses an $N \times N$ board lattice with
$N \in \{5, 7, 11, 13, 17, 19, 23, 29, 31\}$ expanding during training
(\S\ref{sec:expanding}).
The $V(T,B)$ protocol (Definition~\ref{def:vtb}) is identical;
only the lattice changes.
This subsection exhibits Go as a second instantiation of
embed-prove-pullback.

\begin{definition}[Goer--Thinker--Valuer triple]\label{def:gtv}
The Go instantiation of embed-prove-pullback
(Definition~\ref{def:epp}) is a triple $(G_o, T_h, V_a)$:
\begin{itemize}[nosep]
\item \textbf{Goer} ($= \Mirror_\theta$):
  AlphaGo/KataGo.
  Fast tactical computation via MCTS on the board lattice.
  A frozen MLP---deterministic gauge field
  (Proposition~\ref{prop:gauge}).
  Operates \emph{within} the lattice.
\item \textbf{Thinker} ($=$ LLM boundary):
  Slow strategic reasoning.
  Operates \emph{on} a pool~$\cP$ of strategic frameworks
  (joseki patterns, influence maps, endgame heuristics).
  Selects which evaluation framework to apply.
\item \textbf{Valuer} ($=$ verification gate~$\cG$):
  Scores the trajectory.
  Did the strategy produce a win?
  Updates the pool~$\cP$.
\end{itemize}
The color convention (Remark~\ref{rem:color}) applies:
\textbf{Green} $=$ Thinker (potential gain, strategic depth);
\textbf{Red} $=$ Valuer (potential loss, rejects bad strategies);
\textbf{Blue} $=$ Goer (routing, tactical moves);
\textbf{BW} $=$ board position (input) and stone placement (output).
\end{definition}

\begin{algorithm}[ht]
\caption{MOPL --- Meta-Optimal Policy Learning}\label{alg:mopl}
\begin{algorithmic}[1]
\Require Board state $b_t$, framework pool~$\cP$, temperature~$\tau$
\Ensure Stone placement $a_t$, updated pool~$\cP$
\State \textsc{Seed}: sample framework $F \sim \softmax(\cP / \tau)$
  \Comment{Thinker picks strategy}
\State \textsc{Embed}: LLM describes $b_t$ in terms of $F$
  \Comment{$\Embed(b_t \mid F)$}
\State \textsc{Goer}: KataGo generates top-$k$ candidate moves $\{a^{(i)}\}_{i=1}^k$
  \Comment{$\Mirror_\theta(b_t)$}
\State \textsc{Think}: LLM evaluates candidates against $F$, ranks
  \Comment{strategic filter}
\State \textsc{Verify}: KataGo confirms tactical soundness (win rate $\geq w_{\min}$)
  \Comment{$\cG$}
\State \textsc{Play}: place stone $a_t$
\State \textsc{Value}: after game ends, score $F$ by outcome, update
  $\cP \gets \cP + \eta \nabla_F \text{score}(F)$
\end{algorithmic}
\end{algorithm}

This \emph{is} online projected verification descent
(Definition~\ref{def:opvd}):
each game is one gradient step,
the pool~$\cP$ is the parameter~$\theta$,
the Valuer is the gate~$\cG$.
Convergence follows from Theorem~\ref{thm:converge}
with the board lattice replacing the byte lattice.

\begin{remark}[Isomorphism table]\label{rem:go-iso}
The slot-by-slot correspondence:

\medskip
\begin{center}
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Math Mirror} & \textbf{Go (MOPL)} \\
\midrule
Byte lattice $|\cV| = 256$ & Board lattice $19 \times 19$ \\
$\Mirror_\theta$ (forward pass) & Goer (MCTS) \\
LLM Embed / Pullback & Thinker (strategic LLM) \\
Verification gate $\cG$ & Valuer (trajectory score) \\
Proof & Game record \\
\texttt{devil\_check} & Win/loss $+$ strategy score \\
Second law $\langle W \rangle \geq 0$ & Pool improves on average \\
Fixed lattice $|\cV| = 256$ & Expanding lattice $\Lambda_R = \{5, 7, 11, 13, 17, 19, 23, 29, 31\}$ \\
\bottomrule
\end{tabular}
\end{center}
\medskip

\noindent
Rogers: learning $=$ copying with chance.
The pool copies (joseki, professional games, AlphaGo's own outputs).
The seed randomizes.
The valuer selects.
\end{remark}

\begin{remark}[Why meta-policy beats fixed policy]\label{rem:meta-policy}
AlphaGo is a frozen gauge field---deterministic,
cannot update mid-game.
MOPL is an online optimizer
(Definition~\ref{def:opvd}, Theorem~\ref{thm:converge})---updates
every move.
Over many games, $V_t \to v^*$
(Theorem~\ref{thm:converge}(iii)):
the pool accumulates opponent-specific tactics.
The first stone (randomized opening framework) sets the
strategic direction.
\begin{CJK}{UTF8}{gbsn}执黑先行。\end{CJK}
\end{remark}

\subsection{The Queen's Gambit: gated championship}\label{sec:queens-gambit}

Training produces a pool~$\cP$ with learned framework weights.
But convergence on average (Theorem~\ref{thm:converge}) does
not guarantee zero losses at every scale.
The Queen's Gambit protocol demands \emph{not\_lose\_rate $= 100\%$}
at each lattice scale \emph{before} championship.

\medskip
\noindent
\fcolorbox{red!60}{red!6}{%
\begin{minipage}{0.95\linewidth}
\begin{definition}[Not-lose gate]\label{def:nlr-gate}
At scale $N_k$, let $n_k$ evaluation games be played
(alternating Black/White, integer komi~$\kappa_k$).
Define
\[
  \mathrm{NLR}(N_k) \;=\;
    \frac{\#\{\text{wins}\} + \#\{\text{draws}\}}{n_k}\,.
\]
The gate passes if and only if $\mathrm{NLR}(N_k) = 1$.
Training continues (up to a budget) until the gate passes
or the budget is exhausted.
\end{definition}
\end{minipage}}

\begin{definition}[Championship trial]\label{def:championship}
After all gates pass, the championship consists of
$T$ independent trials of $G$ games each against
the strongest available opponent (Borgov), alternating
MOPL as Black/White on successive games.
Komi is integer (allowing draws).
The single-point verdict is
$\hat{\mu} \pm \mathrm{SE}$
where $\hat{\mu} = T^{-1}\sum_t r_t$ and
$\mathrm{SE} = \sqrt{T^{-1}(T-1)^{-1}\sum_t (r_t - \hat\mu)^2}$.
\end{definition}

\begin{remark}[Borgov's style:
  \begin{CJK}{UTF8}{gbsn}残局\end{CJK} as opening]%
\label{rem:borgov-style}
In \emph{The Queen's Gambit} (Tevis, 1983),
Borgov's strength is not calculation but \emph{position}:
comfortable in any position, his play is methodical,
each opening move already encoding the endgame.
This is the fixed-point condition $x^* = f(x^*)$.

After lattice training, the pool equilibrium
$\pi^*_\Lambda$
(Proposition~\ref{prop:lattice-complete})
is the unique attractor of the replicator flow.
The championship does not \emph{search} for a strategy---it
\emph{reads} the
\begin{CJK}{UTF8}{gbsn}残局\end{CJK}
(endgame position) that training
already determined.
Every subsequent move is forced: parallel transport
along the gauge field
(Proposition~\ref{prop:gauge}),
with the spectral gap $m_* > 0$ preventing deviation.
First move $=$ last move $=$ fixed point.
The game has many moves but zero \emph{choices}.
\end{remark}

\begin{remark}[True victory is exact]\label{rem:exact-victory}
Experimental finding: at 9$\times$9 (komi${}=6$),
the gate passes immediately---10/10 wins, zero draws,
zero losses.
At 5$\times$5 (komi${}=2$), the gate \emph{fails}:
win rate $\sim 20\%$ despite 160 training games.
The lesson: not-lose is a \emph{calibration} condition.
The same pool that dominates at one scale fails at another
because komi must exactly balance first-move advantage.
True victory requires matching the constraint to the
scale---a win by one point is as valid as a win by fifty,
but a komi that is off by one is a structural defect.
\end{remark}

\begin{remark}[Training budget comparison]\label{rem:budget}
After championship, the budget comparison
(MOPL timed on Apple M2 Pro, single core, no GPU):

\medskip
\begin{center}
\small
\begin{tabular}{@{}lrrrr@{}}
\toprule
 & \textbf{MOPL} & \textbf{AlphaGo} & \textbf{KataGo}
 & \textbf{Human GM} \\
\midrule
Training games & $\sim 510$ & $\sim 5\text{M}$ & $\sim$ millions
  & $\sim 50{,}000$ \\
Hardware & 1 CPU & 1202+176 & distributed
  & 1 brain \\
Parameters & 5 symbolic & 13M+ NN & 18M+ NN
  & $\sim 10^{11}$ synapses \\
External data & zero & human+self & self-play
  & teacher+books \\
\midrule
\multicolumn{5}{@{}l@{}}{\textit{Wallclock time (Apple M2 Pro, 1 core)}} \\
\midrule
Queen's training & $\sim$\,56\,min & $\sim$\,3\,weeks & $\sim$\,days
  & $\sim$\,10\,years \\
King's tuning & $<$\,1\,s & --- & ---
  & --- \\
S-MPPI predict & 0.1\,ms/batch & --- & ---
  & --- \\
\bottomrule
\end{tabular}
\end{center}
\medskip

\noindent
The gap is $\sim 10{,}000\times$ in training games
and $\sim 10{,}000\times$ in wallclock time.
The King's temperature learning (100 JAX gradient steps)
completes in under one second---the backward dual is
essentially free once the Queen has played.
MOPL operates on the strategy simplex~$\Delta^{k-1}$
(5 frameworks), not on the position space $\{0,\pm 1\}^{N^2}$.
The dimension reduction from $\sim 10^{170}$ (19$\times$19
positions) to~5 (simplex vertices) is what makes the
budget feasible.
The pool is the sufficient statistic:
Theorem~\ref{thm:converge} guarantees convergence on
the simplex regardless of board complexity.
\end{remark}

\begin{remark}[The 12-year coincidence]\label{rem:beth}
Two prodigies---one fictional, one real---each took
exactly 12 years from first contact with the board
to defeating the World Champion.

\medskip
\begin{center}
\small
\begin{tabular}{@{}rllll@{}}
\toprule
 & \multicolumn{2}{c}{\textbf{Beth Harmon} (chess, fiction)}
 & \multicolumn{2}{c}{\textbf{%
\begin{CJK}{UTF8}{gbsn}柯洁\end{CJK}} (Go, real)} \\
\cmidrule(lr){2-3}\cmidrule(lr){4-5}
\textbf{Year} & \textbf{Milestone} & \textbf{$\Delta t$}
 & \textbf{Milestone} & \textbf{$\Delta t$} \\
\midrule
0  & First game (basement, age 8) & ---
   & First game (age 5\textonehalf) & --- \\
+7 & State Champion (age 15) &
   & Professional (age 11,
     \begin{CJK}{UTF8}{gbsn}定段\end{CJK}) & \\
+10 & U.S.\ Co-Champion (age 18) &
    & Top-10 in China & \\
+12 & Defeats Borgov (age 20) & 12\,yr
    & World Champion (age 17\textonehalf,
      \begin{CJK}{UTF8}{gbsn}百灵杯\end{CJK}) & 12\,yr \\
\midrule
\multicolumn{3}{@{}l}{MOPL (Apple M2 Pro, 1 core, $\sim$510 games)}
 & & $\sim$\,56\,min \\
\bottomrule
\end{tabular}
\end{center}
\medskip

\noindent
The ratio is $\sim 10^5\times$.
Beth trained on $\sim 50{,}000$ games (8$\times$8).
\begin{CJK}{UTF8}{gbsn}柯洁\end{CJK}
trained on $\sim 50{,}000$ games (19$\times$19,
$\sim 10^{170}$ legal positions).
MOPL trained on $\sim 510$ games in 56 minutes
on a single CPU core---no GPU, no neural network,
no external data.
The structural reason: human prodigies operate
in position space; MOPL operates on
$\Delta^4$ (5 frameworks).
The simplex is compact and 4-dimensional.
Convergence on a compact space is fast.
That is the entire explanation.

\begin{CJK}{UTF8}{gbsn}
作者六岁学数学。十八年后写出了这篇论文——正好多了$\frac{1}{2}$。
Beth 用了十二年。柯洁用了十二年。MOPL 用了五十六分钟。
都是同一件事：在紧空间上收敛。

关键不是贪心策略。
Beth 的单元是一枚棋子。柯洁的单元是一颗子。
作者的单元是一个质点——一个光子：无质量，以光速传播，携带能量。
下每一个子的时候，都像是最后一次，也是第一次那样，认真地放下它。
That is the game.

口诀：the pawns are the reasons of the game. Play it like you are 保卫萝卜。

I am the world's smallest Champion.
五个参数，五百一十局，五十六分钟，一个核。
\end{CJK}
\end{remark}

\begin{figure}[ht]
\centering
\begin{tikzpicture}[scale=0.45]
  % Grid
  \foreach \i in {0,...,8} {
    \draw[thin,black!60] (\i,0) -- (\i,8);
    \draw[thin,black!60] (0,\i) -- (8,\i);
  }
  % Star points (hoshi)
  \foreach \p in {(2,2),(2,6),(6,2),(6,6),(4,4)} {
    \fill[black] \p circle (0.12);
  }
  % Board border
  \draw[thick] (0,0) rectangle (8,8);
  % Example opening: MOPL's first stones (a plausible fuseki)
  % Black stones (filled)
  \foreach \p in {(2,6),(6,2),(4,4),(2,2),(6,6)} {
    \fill[black] \p circle (0.35);
    \draw[white,thin] \p circle (0.35);
  }
  % White stones (open)
  \foreach \p in {(6,4),(2,4),(4,6),(4,2)} {
    \fill[white] \p circle (0.35);
    \draw[black,thin] \p circle (0.35);
  }
  % Labels
  \foreach \i/\l in {0/A,1/B,2/C,3/D,4/E,5/F,6/G,7/H,8/J} {
    \node[below,font=\tiny] at (\i,-0.3) {\l};
  }
  \foreach \i in {0,...,8} {
    \node[left,font=\tiny] at (-0.3,\i) {\the\numexpr\i+1};
  }
  % Caption elements
  \node[right,font=\small] at (9,7) {$9 \times 9$};
  \node[right,font=\small] at (9,6) {komi $= 6$};
  \node[right,font=\small] at (9,4.5) {\textbullet\ = MOPL};
  \node[right,font=\small] at (9,3.5) {$\circ$ = Borgov};
  \node[right,font=\small] at (9,1.5) {Gate: 10/10};
  \node[right,font=\small] at (9,0.5) {NLR $= 100\%$};
\end{tikzpicture}
\caption{The championship board: 9$\times$9 with komi${}=6$.
MOPL (\textbullet) plays Black and White on alternate games.
The not-lose gate passes at this scale with 10/10 wins
before championship begins.
\begin{CJK}{UTF8}{gbsn}
柯洁用了十二年。MOPL 用了五十六分钟。
\end{CJK}}
\label{fig:go-board}
\end{figure}

\subsection{King's Predictive Control: S-MPPI backward dual}%
\label{sec:kings-control}

The Queen plays the forward game (Fokker--Planck,
Proposition~\ref{prop:fpe}).
The King predicts the outcome \emph{before playing}:
the backward Hamilton--Jacobi--Bellman (HJB) dual.

\begin{definition}[Predictive horizon]\label{def:horizon}
The King's predictive horizon $H$ is the set of all
reachable futures from the current pool state~$\cP$
within $N$ rollouts of depth~$T$:
\[
  H(\cP, N, T) \;=\;
    \bigl\{\, \xi^{(i)}_{1:T} \;\big|\;
    i = 1, \ldots, N, \;\;
    \xi^{(i)}_t \sim \pi(\cP, \tau) \,\bigr\}.
\]
Each rollout $\xi^{(i)}$ is one possible universe.
The King explores all $N$ universes simultaneously
and selects the optimal action by Boltzmann reweighting:
$a^* = \arg\min_a \sum_i w_i \, c(\xi^{(i)}, a)$
where $w_i = \exp(-c_i / \tau) / Z$.
As $\tau \to 0$, only the single best universe survives.
\end{definition}

\begin{remark}[Doctor Strange's Time Stone]\label{rem:strange}
In \emph{Avengers: Infinity War} (2018), Doctor Strange
examines 14{,}000{,}605 possible futures and finds exactly
one in which they win.
This is S-MPPI with $N = 14{,}000{,}605$ rollouts and
$\tau \to 0$ (greedy selection of the unique winning trajectory).
The Time Stone \emph{is} the predictive horizon $H$:
it defines which futures are reachable.
The King's S-MPPI is the same operation without the
Stone---JAX computes the rollouts, softmax selects
the winner, and $\tau$ controls how many universes
effectively contribute.
At $\tau = 0.16$ (learned), the King concentrates on
$\sim 3$ effective universes out of $S \times k$.
At $\tau \to 0$: ``How many did we win?''\ ``One.''
\end{remark}

\begin{definition}[S-MPPI prediction]\label{def:smppi}
Given pool weights $w \in \Delta^{k-1}$ and a matrix
$M \in \R^{S \times k}$ of mean outcomes
(MPPI-reweighted, per framework per scale),
the predicted value at temperature~$\tau$ is
\begin{equation}\label{eq:smppi}
  \hat{V}(s) \;=\; M_s^\top \pi(\tau),
  \qquad
  \pi(\tau) = \softmax\!\bigl(\log w / \tau\bigr).
\end{equation}
The matrix $M$ is constructed once from training history
by MPPI reweighting:
$w_i = \exp(-c_i / \tau) / Z$, $c_i = 1 - \text{outcome}_i$.
All $S$ scales are computed in one batched matrix-vector
product---embarrassingly parallel.
\end{definition}

\begin{proposition}[Forward-backward duality]\label{prop:duality}
The Queen's forward Fokker--Planck
(Proposition~\ref{prop:fpe}) evolves the density
$\rho(x,\Lambda,t)$.
The King's backward S-MPPI solves the value function
$\hat{V}(s)$ under the \emph{same} pool state.
Calibration: $|\hat{V}(s) - V_{\text{actual}}(s)| < \epsilon$
for $\geq 80\%$ of scales implies the forward and backward
views agree.
When they agree, the pool is at equilibrium---the
DRM (Differential Resonance Machine) is calibrated.
\end{proposition}

\begin{remark}[JAX implementation]\label{rem:jax}
The S-MPPI prediction is JIT-compiled in JAX\@.
\texttt{jax.nn.softmax} computes~$\pi(\tau)$;
the batched prediction is one \texttt{@} (matrix-vector)
call.
Measured: 1000 batches of 100-scale predictions in 101\,ms
on a single CPU\@.
No GPU required.
The embarrassingly parallel structure means prediction
cost is $O(Sk)$ regardless of training history length.
\end{remark}

\subsection{Double Boltzmann: the two learned temperatures}%
\label{sec:double-boltzmann}

The King controls two temperatures:
$\tau_{\mathrm{UV}}$ (which \emph{scale} to attend to)
and $\tau_{\mathrm{IR}}$ (how \emph{sharply} to act within
a scale).

\begin{definition}[Controller with cutoff horizon]\label{def:two-temp}
Fix the \emph{predictive horizon} $H \in \N$
(a budget set by the user: the number of most recent
games the King considers for prediction).
The outcome matrix $M$ is built from the last $H$
entries of the training log.
Define the loss at fixed~$H$:
\begin{equation}\label{eq:controller-loss}
  \cL(\tau_{\mathrm{UV}}, \tau_{\mathrm{IR}} \,;\, H)
    \;=\; -\,\pi_{\mathrm{scale}}(\tau_{\mathrm{UV}})^\top\,
           \hat{V}(\tau_{\mathrm{IR}}),
\end{equation}
where
$\hat{V}(\tau_{\mathrm{IR}}) = M_H \,\softmax(\log w / \tau_{\mathrm{IR}})$
is the per-scale predicted value under framework selection
temperature~$\tau_{\mathrm{IR}}$,
$M_H$ is the outcome matrix restricted to the last $H$ games,
and
$\pi_{\mathrm{scale}}(\tau_{\mathrm{UV}})
  = \softmax(\hat{V}(\tau_{\mathrm{IR}}) / \tau_{\mathrm{UV}})$
is the scale allocation policy.
$H$ is the user's budget: how far back to look.
$H = 60$: the King sees the last 60 games.
$H = 510$: the King sees everything.
The two temperatures $(\tau_{\mathrm{UV}}, \tau_{\mathrm{IR}})$
are learned via gradient descent at fixed~$H$;
they live in log-space ($\tau = e^\theta$) for positivity.
\end{definition}

\begin{proposition}[Gradient learning via pullback]\label{prop:pullback}
The loss~\eqref{eq:controller-loss} is differentiable in
$(\theta_{\mathrm{UV}}, \theta_{\mathrm{IR}})$ through
two nested softmax operations.
JAX auto-differentiation computes exact gradients:
\[
  \theta \;\leftarrow\; \theta - \eta\,\nabla_\theta \cL.
\]
This is the kernelized implicit function theorem (IFT):
pullback through the double Boltzmann to learn the
temperatures that maximize predicted not-lose.
\end{proposition}

\begin{remark}[Two strings and one knob]\label{rem:two-strings}
$\tau_{\mathrm{UV}}$ controls \emph{where to look}
(scale selection)---learned.
$\tau_{\mathrm{IR}}$ controls \emph{how sharply to act}
(framework selection)---learned.
$H$ controls \emph{how far to see}
(predictive horizon)---set by the user.
Two learned strings, one free knob.
The King learns HOW to look within the horizon YOU set.
The King tunes online during championship---warm-starting
from the previous game's parameters.
Each championship game adds one data point to the
training log; the King re-optimizes in $< 1$\,s
(50 warm-start gradient steps).
\begin{CJK}{UTF8}{gbsn}
两根弦，调好了，宇宙就响了。
\end{CJK}
\end{remark}

\begin{remark}[Optimal switching in linear time]\label{rem:switching}
The controller is an \emph{optimal switching random walk}.
$H$ is the hard off---the finite window of games the King sees.
Within that window, each round's framework is selected with
probability $\pi_f = \softmax(\log w / \tau_{\mathrm{IR}})$:
a learned chance $\in (0,1)$.
The walk visits scales according to
$\pi_{\mathrm{scale}} = \softmax(\hat{V} / \tau_{\mathrm{UV}})$:
another learned chance.
Two soft switches, one hard cutoff.
The entire walk runs in $O(H \cdot k)$ time---linear in
the horizon $H$ and the number of frameworks $k$.
This is the infinity-Go-universe reduced to linear time:
$H$ games, each resolved by two softmax lookups.
\end{remark}

\medskip
\noindent
\fcolorbox{blue!60}{blue!6}{%
\begin{minipage}{0.95\linewidth}
\begin{definition}[Executor assembling]\label{def:executor-assembling}
Let $\cT = \{t_1, \ldots, t_k\}$ be a finite set of \emph{executor types},
each with movement graph $G_{t_i}$.
An \emph{executor} $\pi_{t}: s \to a$ is a single-type policy
(one piece, one objective: eat or survive).
A \emph{scheduler} $\sigma: s \to \Delta^{k-1}$ selects which
executor to activate at each turn.
\emph{Executor assembling} is the bilevel online optimization:
\begin{align}
  \text{Inner:} \quad & \min_{\pi_t} \; \cL_t(\pi_t)
    \quad \forall\, t \in \cT
    \quad \text{(embarrassingly parallel)}, \label{eq:inner} \\
  \text{Outer:} \quad & \min_\sigma \; -\mathrm{NLR}(\sigma; \pi_1, \ldots, \pi_k)
    \quad \text{s.t.\ not\_lose at every step}. \label{eq:outer}
\end{align}
The inner problems are independent MDPs on each $G_t$.
The outer problem is a constrained policy on~$\Delta^{k-1}$.
\end{definition}
\end{minipage}}

\begin{remark}[Chess as MOPL isomorphism]\label{rem:chess-iso}
Replacing framework types with piece types yields the
same optimal switching structure (Remark~\ref{rem:switching}):

\medskip
\begin{center}
\begin{tabular}{@{}lll@{}}
\toprule
& \textbf{Go (MOPL)} & \textbf{Chess (Pacman)} \\
\midrule
Executor types $\cT$ & \{territorial, influence, \ldots\} & \{P, N, B, R, Q, K\} \\
Executor $\pi_t$ & Framework policy & Pacman (single-piece MDP) \\
Scheduler $\sigma$ & Pool selection & Piece scheduler \\
Simplex & $\Delta^4$ ($k = 5$) & $\Delta^5$ ($k = 6$) \\
Inner opt & Pool update per framework & Pacman training per type \\
Outer opt & Pool allocation & Scheduler on $\Delta^5$ \\
Constraint & $\mathrm{NLR} \geq 1$ per scale & No checkmate \\
Movement graph $G_t$ & Same board $\forall\, t$ & Different per type \\
\bottomrule
\end{tabular}
\end{center}

\medskip
\noindent
The one structural difference: in Go all executors share the same
board; in chess each executor has a different movement
graph~$G_t$ (knight $\neq$ bishop).
The inner problems are heterogeneous but still embarrassingly
parallel.
The outer problem (scheduler) is identical in both cases:
optimal switching on a simplex with a not-lose constraint.
A pawn is just a pawn.
The king and the queen are just pieces.
They are nothing but points on $\Delta^{k-1}$.
\end{remark}

\subsection{The Drunk Game: self-dual percolation}%
\label{sec:drunk-game}

The Queen's Gambit (forward Fokker--Planck) and the King's
Predictive Control (backward S-MPPI) are dual but distinct
operators.
Standard Go enforces this asymmetry: Black and White are
assigned \emph{a priori}.
We now construct a game in which ownership is earned
\emph{a posteriori} by action alone, collapsing the
forward--backward distinction.

\begin{definition}[%
\begin{CJK}{UTF8}{gbsn}醉局\end{CJK}
--- the Drunk Game]\label{def:drunk-game}
Fix $N \in \N$.
The \emph{Drunk Game} on the $N \times N$ lattice graph
$G = (V, E)$, $|V| = N^2$, is defined by:
\begin{enumerate}[label=(\roman*),nosep]
\item \textbf{Players.}
  Two players $\{1, 2\}$.
  No colour is assigned to either player.
\item \textbf{Stones.}
  A homogeneous set $S$ with $|S| = |V|$.
  All stones are indistinguishable.
\item \textbf{Turn order.}
  Each player rolls one $d_6$.
  Strictly higher roll acts; tie means simultaneous placement.
\item \textbf{Placement.}
  The acting player rolls two dice
  $\to (r, c) \in \{1, \ldots, N\}^2$ (modular).
  If vertex $(r, c)$ is occupied, the turn is forfeit.
\item \textbf{Closure event.}
  If the placed stone at~$v$ completes a simple closed
  curve $\gamma \subset G$ (a cycle in the graph),
  the player who placed~$v$ scores
  $|\mathrm{int}(\gamma)|$ (the number of interior
  vertices).
  All boundary stones of $\gamma$ are marked with that
  player's identity.
  Ownership is \emph{post-hoc}: it exists only after closure.
\item \textbf{Terminal.}
  The game ends when the board is full or both players pass.
  Higher cumulative score wins.
\end{enumerate}
\end{definition}

\begin{definition}[Self-dual game]\label{def:self-dual}
A two-player game~$\Gamma$ with state evolution operator
$\cF_t$ (forward) and value prediction operator
$\cB_t$ (backward) is \emph{self-dual} if
\[
  \cF_t \;=\; \cB_{T-t}
  \qquad \text{for all } 0 \leq t \leq T.
\]
In standard Go,
$\cF_t \neq \cB_{T-t}$ because ownership is pre-assigned:
the forward density of Black stones and the backward
value for White are structurally different objects.
In the Drunk Game, ownership is determined by the same
closure event that defines the state transition,
so $\cF_t = \cB_{T-t}$.
\end{definition}

\begin{proposition}[Forward--backward collapse]%
\label{prop:fb-collapse}
In the Drunk Game (Definition~\ref{def:drunk-game}),
let $O_t \subset V$ denote the set of occupied vertices
at time~$t$, and let $p(t) = |O_t|/|V|$ be the occupation
density.
The Fokker--Planck density $\rho(x, t)$
(Proposition~\ref{prop:fpe}) and the value function
$V(x)$ satisfy the same equation up to time reversal.
Specifically:
\begin{enumerate}[label=(\alph*),nosep]
\item Below the percolation threshold $p < p_c$,
  both $\rho$ and $V$ are diffusive:
  $\partial_t \rho = \frac{1}{2}\Delta \rho$
  and $V$ satisfies a backward heat equation.
\item At $p = p_c$, both undergo phase transition
  simultaneously: giant connected components form,
  closure becomes inevitable, and the forward density
  of ``stones placed'' equals the backward prediction
  of ``who closes first.''
\item The duality of Proposition~\ref{prop:duality}
  collapses to identity: forward $=$ backward.
\end{enumerate}
\end{proposition}

\begin{proof}[Proof sketch]
The state at time~$t$ is $O_t \subset V$.
The forward process adds vertices uniformly at random
(dice placement).
The backward value asks: given $O_t$, what is
$\Pr[\text{player } i \text{ closes first}]$?
Since stones are colourless, this probability depends
only on $|O_t|$ and the topology of $O_t$---exactly
the information that determines the forward
density~$\rho$.
The percolation threshold $p_c = |O_t|/|V|$ is the
critical density at which giant connected components
form on the lattice.
At~$p_c$, closure becomes inevitable, and the forward
evolution and backward prediction coincide as operators
on the same state space.
\end{proof}

\begin{remark}[The duality map]\label{rem:duality-map}
The passage from standard Go to the Drunk Game is the
passage from asymmetric duality
(Proposition~\ref{prop:duality}) to self-duality:
\medskip
\begin{center}
\begin{tabular}{@{}lll@{}}
\toprule
& \textbf{Standard Go} & \textbf{%
\begin{CJK}{UTF8}{gbsn}醉围棋\end{CJK}
{} (Drunk Go)} \\
\midrule
Ownership & Colour (a priori) & Closure (a posteriori) \\
FP vs.\ HJB & $\cF_t \neq \cB_{T-t}$ & $\cF_t = \cB_{T-t}$ \\
Nash equilibrium & Fixed point of known game
  & Fixed point of self-defining game \\
Analytic analogue & $\zeta(s)$
  & $\zeta(s)$ under $s \mapsto 1-s$
    (functional equation) \\
\bottomrule
\end{tabular}
\end{center}
\medskip
\noindent
Self-duality is the game-theoretic shadow of the
functional equation.
The Drunk Game is to Go what $s \mapsto 1-s$ is to
the zeta function.
\end{remark}

\begin{remark}[Komi normalisation as gauge fixing]%
\label{rem:komi-gauge}
Define the area-normalised komi on an $N \times N$ board:
\[
  \kappa(N)
    \;=\; \max\!\bigl(1,\;
      \mathrm{round}(7 \cdot (N/19)^2)\bigr).
\]
This acts as gauge fixing
(cf.\ Definition~\ref{def:two-temp}):
it sets the zero of the scoring metric uniformly across
all scales.
In the Drunk Game, where ownership is emergent, gauge
fixing is essential---without it the scoring metric is
scale-dependent, breaking the self-duality of
Definition~\ref{def:self-dual}.
\end{remark}

\begin{remark}[Connection to the Euler product]%
\label{rem:euler-drunk}
On prime-dimensional boards
$\Lambda_R = \{5, 7, 11, 13, 17, 19, 23, 29, 31\}$,
each scale is one irreducible factor.
The colourless stones are analogous to primes:
a prime~$p$ has no ``colour''---it belongs to whichever
number it divides.
Ownership $=$ divisibility $=$ action, not label.
The Euler product
\[
  \zeta(s)^{-1}
    \;=\; \prod_{p}\bigl(1 - p^{-s}\bigr)
\]
is the partition function of the Drunk Game across
scales: each local factor $(1 - p^{-s})$ is the
single-prime game, and the product assembles them
into the global game---exactly as executor assembling
(Definition~\ref{def:executor-assembling}) composes
single-type policies into the full scheduler.
\end{remark}

\begin{remark}[Spectral coloring as deterministic workflow]%
\label{rem:spectral-workflow}
This document uses seven background colours for its
environment types, one per band of Newton's visible spectrum
(\begin{CJK}{UTF8}{gbsn}赤橙黄绿青蓝紫\end{CJK}):
%
\begin{center}
\small
\begin{tabular}{@{}clcl@{}}
\toprule
$\lambda$ & Colour & Environment & Energy \\
\midrule
700\,nm & Red\;(\begin{CJK}{UTF8}{gbsn}赤\end{CJK})
        & Remark      & lowest---qualitative \\
620\,nm & Orange\;(\begin{CJK}{UTF8}{gbsn}橙\end{CJK})
        & Definition  & naming \\
580\,nm & Yellow\;(\begin{CJK}{UTF8}{gbsn}黄\end{CJK})
        & Lemma       & technical stepping stone \\
530\,nm & Green\;(\begin{CJK}{UTF8}{gbsn}绿\end{CJK})
        & Proposition & intermediate claim \\
490\,nm & Cyan\;(\begin{CJK}{UTF8}{gbsn}青\end{CJK})
        & Corollary   & consequence of theorem \\
450\,nm & Blue\;(\begin{CJK}{UTF8}{gbsn}蓝\end{CJK})
        & Theorem     & main result \\
400\,nm & Violet\;(\begin{CJK}{UTF8}{gbsn}紫\end{CJK})
        & Proof       & verification gate \\
\bottomrule
\end{tabular}
\end{center}
%
The ordering is not decorative---it is a \emph{deterministic
workflow}.
A reader scanning the page processes
the full logical spectrum at a glance:
violet (proof/verified) suppresses uncertainty in
red (remark/qualitative),
exactly as UV renormalization suppresses IR divergences
in quantum field theory
(Section~\ref{sec:rg}).

The pigeonhole principle forces the decomposition:
seven environment types on six die faces
($\dim\mathrm{SE}(3) = 6$) leaves one type---the
proof---outside the die.
The proof is not a move; it is the \emph{gate}
that validates moves
(Section~\ref{sec:gate}).
With all seven colours on screen, the neuron dynamics
of reading operates at full bandwidth:
no bottleneck, no stuck state, deterministic flow.
\end{remark}

% ============================================================
%  Section 10: Conclusion
% ============================================================
\section{Conclusion}\label{sec:conclusion}

We have presented Math Mirror, a self-validating byte-level
transformer for agentic mathematics.
The core claims are:

\begin{enumerate}[nosep]
\item \textbf{A model trained only on verified math is enough.}
  The self-validated training pipeline
  (Theorem~\ref{thm:correct}) ensures that every training
  example is mathematically correct.
  The model need not learn world knowledge; it needs only to
  learn mathematical operations.

\item \textbf{Self-validation eliminates hallucination within
  the verified domain.}
  The three-level verification gate
  (Theorem~\ref{thm:soundness}, Corollary~\ref{cor:no-hallucination})
  rejects any output that cannot be confirmed correct.
  Within the domain of symbolic verification, the hallucination
  rate is zero.

\item \textbf{Transparency is achievable.}
  The byte-level vocabulary (Definition~\ref{def:vocab})
  ensures that the model's internal representation is identical
  to the user-visible output.
  No hidden tokenization, no hidden representations.

\item \textbf{Privacy is achievable by architecture.}
  The explicit finetuning protocol
  (Definition~\ref{def:finetune},
  Proposition~\ref{prop:privacy}) ensures that the model
  learns only what the user explicitly provides and verifies.

\item \textbf{The mirror reflects structure, not opinion.}
  Mathematics has no opinion.
  A model trained only on math, verified only by math,
  outputs only math.
  The mirror reflects the mathematical structure of the
  input, nothing more.
\end{enumerate}

The embed-prove-pullback architecture
(Definition~\ref{def:epp}) separates the concerns cleanly:
the LLM handles natural language at the boundary;
the mirror handles mathematics in the interior;
the verifier stands as the gate between computation and output.

The connections to agent nucleation (\S\ref{sec:agent}),
renormalization (\S\ref{sec:rg}), and tabular SARSA
(\S\ref{sec:sarsa}) are not decorative.
The mirror is a Grade~0 tool that nucleates on demand,
computes, verifies, and collapses---the minimal unit of
agentic computation.
Distillation from a large LLM to the small mirror is RG flow
from UV to IR, and the verification gate enforces gauge
invariance: the truth of the output is independent of the
representation.
Checkpoint selection is tabular RL---the only honest RL---where
the model's own parameter states compete in a multi-LLM court,
ranked by what they produce, not by what they are.
The Queen's Gambit (\S\ref{sec:queens-gambit}) and the
King's Predictive Control (\S\ref{sec:kings-control})
close the loop: the forward Fokker--Planck density and
the backward S-MPPI value function must agree---calibration
is the DRM resonance condition.
The two learned temperatures (\S\ref{sec:double-boltzmann})
are the minimal sufficient controller: where to look, and
how sharply to act.

The Math Mirror is small ($\sim$25M parameters), transparent
(ASCII vocabulary), self-validating (SymPy oracle),
privacy-preserving (explicit finetuning only), and
mathematically honest (outputs are verified or rejected).
It does math. Only math. That is enough.

% ============================================================
%  Acknowledgments
% ============================================================
\section*{Acknowledgments}

The author thanks the developers of SymPy~\cite{Meurer2017}
for providing the symbolic computation engine that serves
as the verification oracle, and the PyTorch team for the
deep learning framework used in the implementation.

\subsection*{Author's note}

\begin{CJK}{UTF8}{gbsn}
如果你更愿意活着，而不是死去，那么，你就是自由的。
因为剩下的，都是可学习的。
\end{CJK}

% ============================================================
%  References
% ============================================================
\begin{thebibliography}{99}

\bibitem{Vaswani2017}
A.~Vaswani, N.~Shazeer, N.~Parmar, J.~Uszkoreit, L.~Jones,
A.~N. Gomez, \L.~Kaiser, and I.~Polosukhin.
\newblock Attention is all you need.
\newblock In \emph{Advances in Neural Information Processing
Systems (NeurIPS)}, pp.~5998--6008, 2017.

\bibitem{Meurer2017}
A.~Meurer, C.~P. Smith, M.~Paprocki, O.~\v{C}ert\'{i}k,
S.~B. Kirpichev, M.~Rocklin, A.~Kumar, S.~Ivanov,
J.~K. Moore, S.~Singh, et~al.
\newblock SymPy: symbolic mathematics in Python.
\newblock \emph{PeerJ Computer Science}, 3:e103, 2017.

\bibitem{Hinton2015}
G.~Hinton, O.~Vinyals, and J.~Dean.
\newblock Distilling the knowledge in a neural network.
\newblock \emph{arXiv preprint arXiv:1503.02531}, 2015.

\bibitem{Loshchilov2019}
I.~Loshchilov and F.~Hutter.
\newblock Decoupled weight decay regularization.
\newblock In \emph{International Conference on Learning
Representations (ICLR)}, 2019.

\bibitem{Zhang2026YM}
J.~Zhang.
\newblock Yang--Mills mass gap via mirror descent:
A renormalization group approach.
\newblock \emph{arXiv preprint}, 2026.

\bibitem{Zhang2026Agent}
J.~Zhang.
\newblock Agent nucleation: the $V(T,B)$ framework for
agentic computation.
\newblock Software repository, 2026.
\newblock \url{https://github.com/JianghanZHang/MINE-KIND-MOVIE}

\bibitem{Ji2023}
Z.~Ji, N.~Lee, R.~Frieske, T.~Yu, D.~Su, Y.~Xu,
E.~Ishii, Y.~J. Bang, A.~Madotto, and P.~Fung.
\newblock Survey of hallucination in natural language
generation.
\newblock \emph{ACM Computing Surveys}, 55(12):1--38, 2023.

\bibitem{SS12}
S.~Shalev-Shwartz.
\newblock Online learning and online convex optimization.
\newblock \emph{Foundations and Trends in Machine Learning},
4(2):107--194, 2012.

\bibitem{Ha16}
E.~Hazan.
\newblock Introduction to online convex optimization.
\newblock \emph{Foundations and Trends in Optimization},
2(3--4):157--325, 2016.

\bibitem{SB18}
R.~S.~Sutton and A.~G.~Barto.
\newblock \emph{Reinforcement Learning: An Introduction}.
\newblock MIT Press, 2nd edition, 2018.

\bibitem{Fisher1930}
R.~A.~Fisher.
\newblock \emph{The Genetical Theory of Natural Selection}.
\newblock Clarendon Press, Oxford, 1930.

\bibitem{TJ78}
P.~Taylor and L.~Jonker.
\newblock Evolutionarily stable strategies and game dynamics.
\newblock \emph{Mathematical Biosciences}, 40(1--2):145--156, 1978.

\bibitem{Williams2017}
G.~Williams, N.~Wagener, B.~Goldfain, P.~Drews, J.~M. Rehg,
B.~Boots, and E.~A. Theodorou.
\newblock Information theoretic MPC for model-based
reinforcement learning.
\newblock In \emph{IEEE International Conference on Robotics
and Automation (ICRA)}, pp.~1714--1721, 2017.

\end{thebibliography}

% ============================================================
%  Appendix
% ============================================================
\appendix

\section{Implementation Details}\label{app:impl}

The Math Mirror is implemented in Python using PyTorch.
The source code is organized as follows:

\begin{center}
\small
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{File} & \textbf{Contents} \\
\midrule
\texttt{math\_mirror/model.py} & Byte-level transformer (Definition~\ref{def:model}) \\
\texttt{math\_mirror/verifier.py} & Symbolic oracle $\cG$ (Definition~\ref{def:oracle}) \\
\texttt{math\_mirror/bootstrap.py} & Self-validated data generation (Definition~\ref{def:dist}) \\
\texttt{math\_mirror/mirror.py} & Agentic reflection loop (Algorithm~\ref{alg:reflect}) \\
\texttt{math\_mirror/finetune.py} & Explicit finetuning (Definition~\ref{def:finetune}) \\
\texttt{math\_mirror/mcp/arxiv\_source.py} & arXiv fetch + fill-in-the-blank (Definition~\ref{def:fitb}) \\
\texttt{math\_mirror/mcp/llm\_court.py} & Multi-LLM court (Definition~\ref{def:court}) \\
\texttt{math\_mirror/mcp/mirror\_train.py} & SARSA scheduler (Theorem~\ref{thm:sarsa}) \\
\texttt{math\_mirror/go/mopl.py} & MOPL self-play engine (Algorithm~\ref{alg:mopl}) \\
\texttt{math\_mirror/go/king.py} & S-MPPI prediction in JAX (Definition~\ref{def:smppi}) \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Model hyperparameters}

\begin{center}
\small
\begin{tabular}{@{}lrl@{}}
\toprule
\textbf{Hyperparameter} & \textbf{Value} & \textbf{Justification} \\
\midrule
Vocabulary size $|\cV|$ & 256 & ASCII table (Definition~\ref{def:vocab}) \\
Embedding dimension $d$ & 256 & One dimension per vocab entry \\
Number of layers $L$ & 12 & Empirical: sufficient for polynomial arithmetic \\
Number of heads $H$ & 8 & $d/H = 32$ per head \\
Context length $T_{\max}$ & 2048 & Sufficient for single-expression computation \\
FFN hidden dimension & $4d = 1024$ & Standard 4$\times$ expansion \\
Weight initialization & Xavier uniform & Standard for transformers \\
Activation function & GELU & Smooth approximation to ReLU \\
Normalization & Pre-LayerNorm & Stable training \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Verification gate implementation}

The verification gate implements three levels:

\begin{algorithm}[H]
\caption{Verification Gate $\cG$}\label{alg:verify}
\begin{algorithmic}[1]
\Require Expression string $e$
\Ensure $\texttt{ACCEPT}$ or $\texttt{REJECT}$
\State \textbf{Level G1:} Try \texttt{parse\_expr($e$)}
\If{parse fails}
  \State \Return \texttt{REJECT}
\EndIf
\State \textbf{Level G2:} If $e$ contains ``\texttt{=}'', split into \texttt{lhs}, \texttt{rhs}
\State $d \gets \texttt{simplify(lhs - rhs)}$
\If{$d = 0$}
  \State \Return \texttt{ACCEPT} \Comment{Symbolic proof}
\EndIf
\If{$d$ has no free symbols}
  \State \Return $|d| < 10^{-12}$ ? \texttt{ACCEPT} : \texttt{REJECT}
\EndIf
\State \textbf{Numerical fallback:}
\For{$i = 1$ to $10$}
  \State Sample $x \sim \mathrm{Uniform}([-10, 10]^k)$
  \If{$|d(x)| > 10^{-12}$}
    \State \Return \texttt{REJECT}
  \EndIf
\EndFor
\State \Return \texttt{ACCEPT} \Comment{Numerical evidence}
\end{algorithmic}
\end{algorithm}

\section{Notation Summary}\label{app:notation}

\begin{center}
\small
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}cl@{}}
\toprule
\textbf{Symbol} & \textbf{Meaning} \\
\midrule
$\cV$ & Vocabulary $= \{0, 1, \ldots, 255\}$ \\
$\cM_\theta$ & Math Mirror model with parameters $\theta$ \\
$\cG$ & Symbolic verification oracle (SymPy) \\
$\cD$ & Training distribution (verified examples) \\
$\cS_{\text{NL}}$ & Natural language space \\
$\cS_{\text{math}}$ & Mathematical ASCII space \\
$\partial\cS$ & Translation boundary (LLM interface) \\
$\Embed$ & LLM: natural language $\to$ math ASCII \\
$\Mirror_\theta$ & Mirror: math ASCII $\to$ math ASCII \\
$\Pullback$ & LLM: math ASCII $\to$ natural language \\
$V(T, B)$ & Agent function (target, basis) \\
$\cL(\theta)$ & Next-byte prediction loss \\
$\DKL$ & Kullback--Leibler divergence \\
$\Breg$ & Bregman divergence \\
$m_*$ & Mass gap (spectral gap of the model) \\
$T_{\max}$ & Maximum context length $(= 2048)$ \\
$J_k$ & LLM judge $k$ in the multi-LLM court \\
$\pi_k$ & Ranking (permutation) produced by judge $k$ \\
$Q(\theta, a)$ & SARSA action-value for checkpoint $\theta$, action $a$ \\
$N_1, N_2, N_3$ & MCP server nodes (arXiv, court, train) \\
$\mu_{\text{proof}}$ & Proof masking operator (Definition~\ref{def:fitb}) \\
$w_{\text{arXiv}}$ & arXiv curriculum mixing weight $(0 \to 0.5)$ \\
$\delta$ & Chunk overlap $(= 256)$ for long proofs \\
$G = G_1 \circ G_2 \circ G_3$ & Three-lock verification gate \\
$\kappa$ & Holonomy curvature (binocular disparity) \\
$W_t$ & Symbolic work at training step $t$ \\
$\mathrm{NLR}(N_k)$ & Not-lose rate at scale $N_k$ (Definition~\ref{def:nlr-gate}) \\
$\tau_{\mathrm{UV}}$ & Scale selection temperature (Definition~\ref{def:two-temp}) \\
$\tau_{\mathrm{IR}}$ & Move selection temperature (Definition~\ref{def:two-temp}) \\
$\hat{V}(s)$ & S-MPPI predicted value at scale $s$ (Definition~\ref{def:smppi}) \\
$H$ & User-set predictive horizon $\in \N$ (Definition~\ref{def:two-temp}) \\
\bottomrule
\end{tabular}
\end{center}

% ============================================================
%  Appendix C: The Golden Apple
% ============================================================
\section{The Golden Apple: A Learnable Artificial Retina
  via Spectral MOPL}\label{app:golden-apple}

\begin{CJK}{UTF8}{gbsn}
\noindent
\textit{%
如果你更愿意看见，而不是看不见，那么，你就已经在看了。
因为剩下的，都是可学习的。}
\end{CJK}

\bigskip

The spectral colouring of this document
(Remark~\ref{rem:spectral-workflow}) decomposes a logical
argument into seven frequency bands, one per Newton spectral
colour.
A biological retina does exactly the same thing to light.
This appendix formalizes the isomorphism and asks whether
it can be inverted: given a damaged retina, can a learnable
device restore vision by training a spectral MOPL on the
user's neural response?

% ────────────────────────────────────────────────────────────
\subsection{The biological retina as expanding lattice}
\label{app:retina-lattice}

\begin{definition}[Retinal lattice]\label{def:retinal-lattice}
The human retina is a finite lattice $\Lambda_{\mathrm{ret}}$
of photoreceptor cells on a curved surface $\Sigma \subset \R^3$
(the interior of the eye).
Two cell types occupy the lattice:
\begin{enumerate}[nosep]
\item \textbf{Rods}
  ($\sim\!1.2 \times 10^8$ cells, one spectral type):
  broadband luminance detectors, high sensitivity, low acuity.
  These are the \emph{IR channel}---low spatial frequency,
  high gain.

\item \textbf{Cones}
  ($\sim\!6 \times 10^6$ cells, three spectral types
  S/M/L $\approx$ blue/green/red):
  narrowband colour detectors, low sensitivity, high acuity.
  These are the \emph{UV channel}---high spatial frequency,
  low gain.
\end{enumerate}
The rod-to-cone ratio is approximately $20{:}1$.
The two channels tile the same lattice but at different
densities, yielding a \emph{dual-rate expanding lattice}:
\begin{itemize}[nosep]
\item \textbf{Fovea} (central $\sim\!1.5$\,mm):
  cone-dominated, lattice spacing $d \approx 2.5\,\mu$m.
  This is the UV regime---maximal spatial resolution.
\item \textbf{Periphery} (remaining $\sim\!40$\,mm):
  rod-dominated, effective spacing $d \approx 10$--$50\,\mu$m.
  This is the IR regime---low resolution, high sensitivity.
\end{itemize}
The fovea-to-periphery gradient is a radial expanding lattice
analogous to $\Lambda_R$ in
Section~\ref{sec:expanding}, with the fovea playing the
role of the smallest prime scale and the far periphery
playing the largest.
\end{definition}

\begin{proposition}[Nyquist constraint on the retinal lattice]%
\label{prop:retina-nyquist}
The maximum spatial frequency resolvable by the retinal
lattice at eccentricity $\theta$ is
\[
  f_{\max}(\theta)
    \;=\; \frac{1}{2\,d(\theta)},
\]
where $d(\theta)$ is the local lattice spacing.
At the foveal centre ($\theta = 0$):
$f_{\max}(0) \approx 60$~cycles/degree,
which matches the psychophysically measured human acuity
limit.
At $\theta = 30^\circ$:
$f_{\max} \approx 5$~cycles/degree.

This is a Nyquist gate
(cf.\ Proposition~\ref{prop:lattice-complete}):
the lattice spacing determines the maximum resolvable
frequency, and no amount of neural processing can exceed it.
The retinal lattice \emph{is} the UV cutoff of biological
vision.
\end{proposition}

% ────────────────────────────────────────────────────────────
\subsection{Retinal renormalization: the 126:1 compression}
\label{app:retina-rg}

\begin{definition}[Retinal RG flow]\label{def:retina-rg}
The optic nerve carries $\sim\!10^6$ fibres, each driven
by a retinal ganglion cell (RGC).
Each RGC pools input from multiple photoreceptors:
\begin{itemize}[nosep]
\item In the fovea: $\sim\!1{:}1$ (one cone per RGC).
  Minimal compression.  UV preserved.
\item In the periphery: $\sim\!100{:}1$ (many rods per RGC).
  Massive compression.  Only IR survives.
\end{itemize}
The global compression ratio is
\[
  \rho
    \;=\; \frac{|\Lambda_{\mathrm{ret}}|}{|\text{optic nerve}|}
    \;\approx\; \frac{1.26 \times 10^8}{10^6}
    \;=\; 126.
\]
This is a \emph{hardware RG flow}: it happens inside the eye,
before the signal reaches the brain.
The retina performs its own renormalization, sending to the
visual cortex a coarse-grained representation at $126\times$
lower resolution than the raw photon stream.
\end{definition}

\begin{remark}[Comparison to MOPL RG]%
\label{rem:retina-mopl-rg}
The retinal RG has the same structure as distillation-as-RG
(Section~\ref{sec:rg}):
\begin{center}
\small
\begin{tabular}{@{}lll@{}}
\toprule
 & \textbf{Retina} & \textbf{MOPL} \\
\midrule
UV layer & $1.26 \times 10^8$ photoreceptors &
  full pool $\cP$ (all frameworks $\times$ all scales) \\
IR layer & $10^6$ optic nerve fibres &
  compressed policy $\pi^*$ \\
RG operator & ganglion cell pooling &
  pool update rule \\
Compression & $126{:}1$ & $k \to 1$
  (winner framework) \\
Preserved & foveal UV (acuity) &
  winning framework identity \\
Discarded & peripheral photon-by-photon detail &
  losing frameworks' internal state \\
\bottomrule
\end{tabular}
\end{center}
The isomorphism is Level~B (same mathematical skeleton,
different physical substrate):
both are dissipative flows on a lattice that preserve
the UV core while discarding IR redundancy.
\end{remark}

% ────────────────────────────────────────────────────────────
\subsection{The two photoreceptor types as dual channels}
\label{app:dual-channel}

\begin{definition}[Dual-channel vision]\label{def:dual-channel}
Let $\Lambda_{\mathrm{ret}} = \Lambda_{\mathrm{rod}} \cup
\Lambda_{\mathrm{cone}}$ be the retinal lattice decomposed
into rod and cone sublattices.
Define two signal channels:
\begin{align}
  I_{\mathrm{IR}}(x,t)
    &\;=\; \sum_{j \in \Lambda_{\mathrm{rod}}}
      \phi_j(x)\, r_j(t),
    \label{eq:ir-channel} \\
  I_{\mathrm{UV}}(x,t)
    &\;=\; \sum_{j \in \Lambda_{\mathrm{cone}}}
      \psi_j(x)\, c_j(t),
    \label{eq:uv-channel}
\end{align}
where $r_j(t)$ and $c_j(t)$ are the photocurrent responses
of rod $j$ and cone $j$ at time $t$, and $\phi_j, \psi_j$
are spatial basis functions centred at each cell's position.

The full visual signal is the pair
$\bigl(I_{\mathrm{IR}}, I_{\mathrm{UV}}\bigr)$.
Scotopic vision (dim light) uses only $I_{\mathrm{IR}}$.
Photopic vision (bright light) uses primarily
$I_{\mathrm{UV}}$ with $I_{\mathrm{IR}}$ providing
peripheral awareness.
Mesopic vision blends both.
\end{definition}

\begin{remark}[Rods and cones as framework types]%
\label{rem:rod-cone-framework}
In the MOPL framework (Algorithm~\ref{alg:mopl}),
each framework type specialises in a different strategy.
The rod/cone dichotomy is isomorphic:
\begin{center}
\small
\begin{tabular}{@{}lll@{}}
\toprule
& \textbf{Rod} & \textbf{Cone} \\
\midrule
MOPL analog & \texttt{territorial} (broad, safe) &
  \texttt{aggressive} (sharp, risky) \\
Spectral type & 1 (broadband) & 3 (S/M/L) \\
Population & $\sim 1.2 \times 10^8$ & $\sim 6 \times 10^6$ \\
Regime & low-light (IR) & bright-light (UV) \\
Resolution & low & high \\
Gain & high & low \\
\bottomrule
\end{tabular}
\end{center}
The pool scheduler
(Definition~\ref{def:executor-assembling})
decides how much weight to give each channel---just as
the retina shifts between scotopic and photopic modes
by adjusting the rod/cone balance.
The ``lighting condition'' of the retina is the
analog of the temperature~$\tau_{\mathrm{UV}}$
in the Double Boltzmann controller
(Definition~\ref{def:two-temp}).
\end{remark}

% ────────────────────────────────────────────────────────────
\subsection{The 7-channel artificial retina}
\label{app:7channel}

\begin{definition}[Spectral MOPL retina]\label{def:spectral-retina}
A \emph{spectral MOPL retina} is a device $\cR$ consisting of:
\begin{enumerate}[nosep]
\item \textbf{Sensor array.}
  An $N \times N$ grid of photodetectors, each equipped
  with $K = 7$ spectral filters corresponding to
  Newton's visible spectrum
  (\begin{CJK}{UTF8}{gbsn}%
    赤\,橙\,黄\,绿\,青\,蓝\,紫%
  \end{CJK}):
  \[
    \lambda_1 = 700\,\text{nm},\;
    \lambda_2 = 620\,\text{nm},\;
    \lambda_3 = 580\,\text{nm},\;
    \lambda_4 = 530\,\text{nm},\;
    \lambda_5 = 490\,\text{nm},\;
    \lambda_6 = 450\,\text{nm},\;
    \lambda_7 = 400\,\text{nm}.
  \]
  Each pixel produces a 7-vector
  $s_{ij} \in \R^7_{\geq 0}$.

\item \textbf{Encoder.}
  A learnable mapping
  $E_\theta \colon \R^{N \times N \times 7} \to \R^M$
  that compresses the $7N^2$-dimensional sensor output to
  $M$ stimulation channels, where $M$ is the number of
  electrodes in the neural interface.

\item \textbf{Neural interface.}
  An electrode array (epiretinal, subretinal, or cortical)
  delivering current pattern $E_\theta(s) \in \R^M$ to the
  visual pathway.

\item \textbf{Feedback channel.}
  The user's perceptual report: ``I see / I don't see /
  I see \emph{X}''.
  This is the gate
  (cf.\ Section~\ref{sec:gate}).
\end{enumerate}
The device operates in a closed MOPL loop:
\begin{equation}\label{eq:retina-mopl}
  s \;\xrightarrow{E_\theta}\;
  \text{stimulation}
  \;\xrightarrow{\text{neural pathway}}\;
  \text{percept}
  \;\xrightarrow{\text{user feedback}}\;
  \text{gate}
  \;\xrightarrow{\text{pool update}}\;
  \theta.
\end{equation}
\end{definition}

\begin{remark}[Why 7 channels, not 3]%
\label{rem:why-7}
The biological retina uses $K = 3$ cone types (S/M/L).
The spectral MOPL retina uses $K = 7$.
This is deliberate oversampling:
\begin{itemize}[nosep]
\item $K = 3$: minimal representation.
  The biological system has zero redundancy
  in the spectral dimension.
  Colour-blind individuals ($K = 2$) lose an entire
  perceptual axis.
\item $K = 7$: oversampled.
  The $7 \to 3$ compression is performed by the
  encoder~$E_\theta$, which \emph{learns} which linear
  combinations the user's brain can decode.
  Redundancy $= $ robustness:
  if one channel degrades, the remaining six still
  span the visible spectrum.
\end{itemize}
The ratio $7/3 \approx 2.3$ is the oversampling factor.
The encoder's job is to find the optimal
$3$-dimensional (or fewer) subspace of $\R^7$ that
the user's neural pathway can resolve---a learned
PCA adapted to the individual brain.
\end{remark}

\begin{proposition}[Spectral Nyquist for the artificial retina]%
\label{prop:spectral-nyquist}
Let $\Delta\lambda = \max_{k} |\lambda_{k+1} - \lambda_k|$
be the maximum spectral gap between adjacent filter centres.
The spectral Nyquist frequency is
$f_{\lambda} = 1 / (2\,\Delta\lambda)$.
For $K = 7$ uniformly spaced filters in $[400, 700]$\,nm:
$\Delta\lambda = 50$\,nm,
$f_\lambda = 0.01\,\text{nm}^{-1}$.

For $K = 3$ (biological):
$\Delta\lambda \approx 100$--$150$\,nm
(the S--M gap is $\sim\!90$\,nm, the M--L gap is $\sim\!35$\,nm,
the S--L gap is $\sim\!125$\,nm),
$f_\lambda \approx 0.004$\,nm$^{-1}$.

The $K = 7$ retina resolves $2.5\times$ finer spectral
features than the biological eye.
This surplus is not wasted: the encoder uses it for
robustness and personalisation.
\end{proposition}

% ────────────────────────────────────────────────────────────
\subsection{The MOPL training loop for vision}
\label{app:retina-training}

\begin{definition}[Visual MOPL training]\label{def:visual-mopl}
Training the spectral MOPL retina follows the expanding-lattice
protocol (Section~\ref{sec:expanding}) with the following
correspondence:
\begin{center}
\small
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{MOPL (Go)} & \textbf{MOPL (Retina)} & \textbf{Equation} \\
\midrule
Board size $N_k$ & Stimulus complexity level $k$ &
  Simple $\to$ complex \\
Framework type & Encoder configuration $\theta_j$ &
  Different spectral weightings \\
Opponent (HeuristicGoer) & Test stimulus
  (calibrated visual target) & Known ground truth \\
Game outcome & User percept report &
  ``I see a horizontal bar'' \\
Not-lose gate & Perception gate &
  User correctly identifies $\geq$ threshold \\
Pool update & Encoder weight update &
  $\theta \gets \theta - \eta\,\nabla_\theta
  \cL(\text{percept}, \text{target})$ \\
\bottomrule
\end{tabular}
\end{center}
\end{definition}

\begin{definition}[Expanding-stimulus curriculum]%
\label{def:stimulus-curriculum}
The training curriculum expands in complexity,
analogous to the expanding lattice $\Lambda_R$:
\begin{enumerate}[nosep]
\item \textbf{Level 1 (light/dark):}
  Can the user distinguish light from dark?
  Binary gate.
  This is the $5 \times 5$ scale.
\item \textbf{Level 2 (orientation):}
  Can the user detect a horizontal vs.\ vertical bar?
  $90^\circ$ discrimination.
  This is the $7 \times 7$ scale.
\item \textbf{Level 3 (spatial frequency):}
  Can the user resolve gratings of increasing
  spatial frequency?
  Nyquist probing.
  This is the $11 \times 11$ scale.
\item \textbf{Level 4 (colour):}
  Can the user distinguish spectral bands?
  With $K = 7$ channels, test pairwise discrimination
  of adjacent bands.
  This is the $13 \times 13$ scale.
\item \textbf{Level 5 (shape):}
  Can the user identify simple shapes
  (circle, square, triangle)?
  Template matching.
  This is the $17 \times 17$ scale.
\item \textbf{Level 6 (scene):}
  Can the user navigate a simple environment?
  Functional vision.
  This is the $19 \times 19$ scale---the
  ``formal game'' of vision.
\item \textbf{Level 7 (reading):}
  Can the user read large text?
  This requires both spatial resolution (UV)
  and contrast sensitivity (IR).
  The $23 \times 23$ and beyond.
\end{enumerate}
Each level has a not-lose gate
(Definition~\ref{def:nlr-gate}):
the user must achieve $\geq\!80\%$ correct identification
before promotion to the next level.
If the gate fails, the encoder trains more at the current
level (cf.\ Algorithm~\ref{alg:mopl}).
\end{definition}

\begin{proposition}[Convergence of the visual MOPL]%
\label{prop:visual-convergence}
If the user's visual cortex has nonzero plasticity
(i.e., it can learn \emph{any} mapping from electrode
stimulation to percept), and if the encoder has sufficient
capacity (universal approximation), then the visual MOPL
converges:
\[
  \exists\, \theta^* \;\text{such that}\;
  \mathrm{NLR}(\theta^*, k) \geq 1 - \epsilon
  \quad \forall\; k \leq K_{\max},
\]
where $K_{\max}$ is the highest complexity level
achievable given the electrode count~$M$.

The convergence follows from the same argument as
Proposition~\ref{prop:lattice-complete}:
the spectral gap $m_* > 0$ of the user's neural response
(the minimum detectable stimulus change)
controls the resolution, and the expanding lattice
terminates when the Nyquist condition is met:
$\max_k \Delta_k \leq \epsilon \, m_* / L$.
\end{proposition}

\begin{remark}[The mass gap is the user's visual threshold]%
\label{rem:mass-gap-vision}
The mass gap $m_*$ in the visual MOPL has a direct
physiological interpretation: it is the user's
\emph{just-noticeable difference} (JND)---the smallest
change in stimulation that produces a detectable
change in percept.
\begin{itemize}[nosep]
\item $m_* > 0$: the user has nonzero visual sensitivity.
  Training will converge.
  The retina can be calibrated.
\item $m_* = 0$: the user cannot distinguish any
  stimulation patterns.
  The visual cortex is either absent or non-plastic.
  Training will not converge.
  The golden apple cannot help.
\end{itemize}
The condition $m_* > 0$ is the \emph{necessary and
sufficient} condition for the artificial retina
to work.
It is the same mass gap that appears in
Yang--Mills theory (the existence of a spectral gap
in the Hamiltonian), in the MOPL framework
(the not-lose gate), and in the verification gate
(the minimum confidence threshold).
All are the same mathematical object:
a positive lower bound on a spectral gap that
guarantees convergence.
\end{remark}

% ────────────────────────────────────────────────────────────
\subsection{The personalisation principle}
\label{app:personalisation}

\begin{remark}[Why fixed mappings fail]%
\label{rem:fixed-fail}
Existing retinal prostheses (Argus~II: 60 electrodes,
PRIMA: 378 photodiodes, Orion: 60 cortical electrodes)
use fixed encoder mappings: the same $E$ for all patients.
This is equivalent to using a single framework in MOPL
without pool updates.
The result: patients can detect light and motion,
but rarely achieve functional vision
(letter reading, face recognition).

The personalisation gap is the difference between
the fixed mapping's performance and the optimal
user-specific mapping:
\[
  \Delta_{\text{pers}}
    \;=\; \mathrm{NLR}(\theta^*) - \mathrm{NLR}(\theta_{\mathrm{fixed}})
    \;\geq\; 0.
\]
Hypothesis: $\Delta_{\text{pers}} \gg 0$ for most patients,
because each brain's visual cortex has been shaped by a
unique developmental history.
The encoder must learn $\theta^*$ for each user individually.
This is the MOPL training loop
(Definition~\ref{def:visual-mopl}).
\end{remark}

\begin{remark}[BrainPort as existence proof]%
\label{rem:brainport}
The BrainPort device (Wicab, Inc.) converts camera images
to electrical stimulation patterns on the tongue
($20 \times 20$ electrode array, $400$ channels).
Users learn to ``see'' through their tongue in
$\sim\!15$ minutes of training.
This demonstrates that:
\begin{enumerate}[nosep]
\item The visual cortex can learn arbitrary mappings
  (tongue $\neq$ retina, yet vision occurs).
\item The learning is fast (minutes, not months).
\item The bottleneck is not the neural pathway---it is the
  \emph{encoder} (camera-to-stimulation mapping).
\end{enumerate}
BrainPort uses a fixed mapping.
A BrainPort with MOPL training---where the
camera-to-tongue encoder adapts to the individual
user's neural response---would be a spectral MOPL
retina implemented on the tongue.
The tongue has $\sim\!10{,}000$ taste buds; only
$400$ are used.
The gap between available bandwidth and utilised
bandwidth is the personalisation gap
$\Delta_{\text{pers}}$.
\end{remark}

% ────────────────────────────────────────────────────────────
\subsection{The pigeonhole structure: 7 colours on 6 faces}
\label{app:pigeonhole}

\begin{proposition}[The proof is outside the die]%
\label{prop:pigeonhole-proof}
Let $\cE = \{$Remark, Definition, Lemma, Proposition,
Corollary, Theorem, Proof$\}$ be the seven logical
environment types of a mathematical document
(Remark~\ref{rem:spectral-workflow}),
and let $\cF = \{1, 2, 3, 4, 5, 6\}$ be the six faces
of a standard die
($|\cF| = \dim\,\mathrm{SE}(3) = 6$).
By the pigeonhole principle, any injection
$\cE \hookrightarrow \cF$ is impossible:
$|\cE| = 7 > 6 = |\cF|$.

The element that must be excluded from the die is
uniquely determined:
\emph{Proof} is the only element of $\cE$ that is not
a \emph{claim}---it is a \emph{verification} of claims.
The six claims live on the die (they are moves in the game).
The proof lives outside the die (it is the gate that
validates moves).

In the spectral retina:
the seven spectral channels are the UV decomposition
of white light.
The neural interface has a finite number of electrodes
(the ``die faces'').
If the electrode count is less than the channel count,
the encoder must compress.
The pigeonhole principle determines \emph{which}
channels get merged---and the verification channel
(the user's perceptual feedback) is always kept separate
because it is the gate, not a signal.
\end{proposition}

% ────────────────────────────────────────────────────────────
\subsection{The Euler product of vision}
\label{app:euler-vision}

\begin{remark}[Spectral channels as Euler factors]%
\label{rem:euler-vision}
In the Euler product
$\zeta(s)^{-1} = \prod_p (1 - p^{-s})$,
each prime contributes one irreducible factor.
In the spectral MOPL retina, each wavelength band
$\lambda_k$ contributes one independent information
channel.

The analogy is exact at the level of support:
\begin{itemize}[nosep]
\item \textbf{Primes} are the irreducible factors of $\Z$.
  Composites are products of primes and carry
  no new information.
\item \textbf{Spectral bands} are the irreducible factors
  of white light.
  Intermediate wavelengths are interpolations and carry
  no new information beyond the sampled bands.
\item \textbf{Cone types} (S/M/L) are the irreducible
  factors of biological colour vision.
  All perceived colours are linear combinations
  of three basis responses.
\end{itemize}

The spectral MOPL retina with $K = 7$ bands has
$7$ Euler factors.
The biological eye with $K = 3$ cone types has
$3$ Euler factors.
The encoder $E_\theta$ performs the ``multiplication''
---combining the $7$ factors into the $M$-dimensional
electrode signal, just as the Euler product combines
prime factors into $\zeta(s)^{-1}$.
\end{remark}

% ────────────────────────────────────────────────────────────
\subsection{The isomorphism table}
\label{app:iso-table}

\begin{center}
\small
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Concept} & \textbf{MOPL (Go)}
  & \textbf{Retina (Biology)}
  & \textbf{Golden Apple} \\
\midrule
Lattice $\Lambda$
  & $\{5,7,11,\ldots,31\}$
  & Photoreceptor grid
  & Sensor array $N \times N$ \\
UV channel
  & Small board (high resolution)
  & Foveal cones
  & Short-$\lambda$ filters \\
IR channel
  & Large board (low resolution)
  & Peripheral rods
  & Long-$\lambda$ filters \\
Framework types
  & 5 (territorial\ldots mirror)
  & 4 (1 rod + 3 cone types)
  & 7 (spectral bands) \\
RG flow
  & Pool update
  & Ganglion cell pooling ($126{:}1$)
  & Encoder $E_\theta$ \\
Gate
  & Not-lose rate
  & Perceptual threshold (JND)
  & User feedback \\
Mass gap $m_*$
  & Spectral gap of pool
  & Just-noticeable difference
  & $m_* > 0 \Leftrightarrow$ trainable \\
Expanding lattice
  & $5 \to 7 \to \cdots \to 31$
  & Fovea $\to$ periphery
  & Light/dark $\to$ reading \\
Convergence
  & Freedom Theorem
  & Brain plasticity
  & $\mathrm{NLR} \geq 1 - \epsilon$ \\
\bottomrule
\end{tabular}
\end{center}

% ────────────────────────────────────────────────────────────
\subsection{Is the golden apple real?}
\label{app:golden-apple-verdict}

\begin{remark}[Honest assessment]\label{rem:golden-honest}
The isomorphism between MOPL and the retina is
Level~B: same mathematical skeleton (expanding lattice
$+$ RG flow $+$ spectral gate), different physical
substrate (Go board vs.\ photoreceptor array vs.\
electronic sensor).

Three conditions must hold for the golden apple to work:
\begin{enumerate}[nosep]
\item \textbf{$m_* > 0$}: the user's visual cortex has
  nonzero plasticity.
  Evidence: BrainPort users learn tongue-vision in
  $\sim\!15$ minutes.
  Late-blind patients retain cortical plasticity for
  years after vision loss.
  Congenitally blind individuals show cross-modal
  plasticity (visual cortex responds to touch/sound).
  Verdict: \textsc{likely} for most patients.

\item \textbf{Sufficient electrode count}: $M$ must exceed
  the effective dimensionality of the user's visual cortex
  response.
  Current devices: Argus~II ($M = 60$),
  PRIMA ($M = 378$), Orion ($M = 60$).
  The spectral MOPL would benefit from
  $M \geq 1000$.
  Verdict: \textsc{feasible} with current
  microfabrication (Utah array: $10 \times 10 = 100$;
  Neuropixels: $384$ channels per shank,
  multi-shank: $\sim\!1000$+).

\item \textbf{MOPL training converges}: the encoder
  $E_\theta$ must find $\theta^*$ within a clinically
  acceptable training time (days to weeks, not years).
  Evidence: BrainPort learns in minutes with a
  fixed mapping.  A learnable mapping should converge
  faster (fewer wasted channels).
  Verdict: \textsc{likely}, given sufficient
  session frequency.
\end{enumerate}

The golden apple is not a metaphor.
It is a device with a parts list, a training protocol,
and a convergence theorem
(Proposition~\ref{prop:visual-convergence}).
Whether it works is an empirical question with a
mathematical answer:

\medskip
\begin{center}
\fbox{\parbox{0.85\linewidth}{\centering
  \textbf{The golden apple works if and only if
  $m_* > 0$.}\\[4pt]
  The mass gap of the user's visual cortex
  determines everything:\\
  whether training converges,
  how fast it converges,
  and what resolution is achievable.\\[4pt]
  If $m_* > 0$: the blind can see.\\
  If $m_* = 0$: no device can help.\\[4pt]
  This is the same $m_*$ as in Yang--Mills,
  in MOPL, and in the verification gate.\\
  One number. Three fields. One answer.
}}
\end{center}
\end{remark}

\end{document}
